---
title: "Junk Code"
output: html_document
---




*We will figure that out later. *
```{r}
CAPE_m5 <- lmer(maxCAPE ~ SSTgradient_30day + avg_Nino3430day + NAO_30day + PNA_30day +  MJO_30day + Lat + Lon+ (1|Mo), data = BigDays.sfdfT)
summary(CAPE_m5)
ranef(CAPE_m5)
```

```{r}
CAPE_m6 <- lmer(maxCAPE ~ SSTgradient_5day + avg_Nino345day + NAO_5day + PNA_5day +  MJO_5day + Lat + Lon+ (1|Mo), data = BigDays.sfdfT)
summary(CAPE_m6)
ranef(CAPE_m6)
```

Compare the effect size and t values between the 5, 10, 20, and 30 day climate influencers. Select the most important ones. 

```{r}
CAPE_m7 <- lmer(maxCAPE ~ SSTgradient_5day + avg_Nino345day + NAO_5day + PNA_5day +  MJO_10day + Lat + Lon+ (1|Mo), data = BigDays.sfdfT)
summary(CAPE_m7)
ranef(CAPE_m7) 
```

```{r}
CAPE_m11 <- lmer(maxCAPE ~ SSTgradient_5day + avg_Nino345day + PNA_5day +  MJO_10day + Lon+ (1|Mo), data = BigDays.sfdfT)
summary(CAPE_m11)
```


```{r}
CAPE_m8 <- lmer(maxCAPE ~  SSTgradient_5day + avg_Nino3420day +  PNA_5day +  MJO_10day + Lon + (1|Mo), data = BigDays.sfdfT)
summary(CAPE_m8)
ranef(CAPE_m8) 
```



```{r}
CAPE_m9<- lmer(maxCAPE ~ SSTgradient_5day +  PNA_5day +  MJO_10day + Lon + (1|Mo), data = BigDays.sfdfT)
summary(CAPE_m9)

```

```{r}
CAPE_m10 <- lm(maxCAPE ~ SSTgradient_5day +  PNA_5day +  MJO_10day + Lon, data = BigDays.sfdfT)
summary(CAPE_m10)
```

```{r}
CAPE_m12 <- lm(maxCAPE ~ Month + SSTgradient_5day +  PNA_5day +  MJO_10day + Lon, data = BigDays.sfdfT)
summary(CAPE_m12)
```

```{r}
CAPE_m13 <- lm(maxCAPE ~ Month +  MJO_10day, data = BigDays.sfdfT)
summary(CAPE_m13)
```

```{r}
AIC(CAPE_m1,CAPE_m2,CAPE_m3,CAPE_m4,CAPE_m5,CAPE_m6,CAPE_m7,CAPE_m8, CAPE_m9,CAPE_m10,CAPE_m11, CAPE_m12, CAPE_m13)
```
Using AIC It would have me believe that model 7 is significantly better than all of the others. 

Is there a relationship between SST 20 day and Nino5day or NAO5day? 

```{r}
cor(BigDays.sfdfT$SSTgradient_20day, BigDays.sfdfT$avg_Nino345day)
# 0.08557304
cor(BigDays.sfdfT$SSTgradient_20day, BigDays.sfdfT$NAO_5day)
#0.0008885289
```

##SLBS: 
*What is the relationship between these variables and SLBS? *
```{r}
SLBS_env <- data.frame(AirTemperature = BigDays.sfdfT$avg_globalairtemp, SSTgradient = BigDays.sfdfT$SSTgradient, ElNino = BigDays.sfdfT$avg_Nino3410day, NAO = BigDays.sfdfT$NAO_10day, PNA = BigDays.sfdfT$PNA_10day, MJO = BigDays.sfdfT$MJO_10day, SLBS = BigDays.sfdfT$maxBS_shallow)


corrplot.mixed(cor(SLBS_env), tl.col="black", lower.col="black", number.cex = 1, tl.cex = 1)
```

```{r}
SLBS_m1 <- lm(maxBS_shallow ~ SSTgradient + avg_Nino3410day + NAO_10day + PNA_10day +  MJO_10day, data = BigDays.sfdfT)
SLBS_m1 <- lmer(maxBS_shallow ~ SSTgradient + avg_Nino3410day + NAO_10day + PNA_10day +  MJO_10day + (1|Mo), data = BigDays.sfdfT)
summary(SLBS_m1)
ranef(SLBS_m1) #There is a strong seasonal variation in SLBS values. 
```

```{r}
m1 <- lm(maxBS_shallow ~ SSTgradient + PNA_10day +  MJO_10day, data = BigDays.sfdfT)
SLBS_m2 <- lmer(maxBS_shallow ~ SSTgradient + PNA_10day +  MJO_10day + (1|Mo), data = BigDays.sfdfT)
summary(SLBS_m2)
ranef(SLBS_m2) 
```

Month is the only random effect. 
```{r}
SLBS_m3 <- lmer(maxBS_shallow ~ Year + SSTgradient + avg_Nino3410day + NAO_10day + PNA_10day +  MJO_10day  + (1|Mo), data = BigDays.sfdfT)
summary(SLBS_m3)
ranef(SLBS_m3)
```

```{r}
SLBS_m4 <- lmer(maxBS_shallow ~ Year + SSTgradient + NAO_10day + PNA_10day +  MJO_10day  + (1|Mo), data = BigDays.sfdfT)
summary(SLBS_m4)
ranef(SLBS_m4)
```

```{r}
SLBS_m5 <- lmer(maxBS_shallow ~ Year + SSTgradient + PNA_10day +  MJO_10day  + (1|Mo), data = BigDays.sfdfT)
summary(SLBS_m5)
ranef(SLBS_m5)
```

*Model 5 is the best*
```{r}
AIC(SLBS_m1 ,SLBS_m2 ,SLBS_m3 ,SLBS_m4 ,SLBS_m5)
```

*Check the residuals of model 5 of SLBS.*
```{r}
plot(residuals(SLBS_m5)) 
```
There is no relationship between the residuals.

*Make a plot of the standardized residuals for the SLBS model*
```{r}
SLBSmod.df <- fortify.merMod(model=SLBS_m5, data = getData(SLBS_m5))
SLBSmod.df$predSLBS <- predict(SLBS_m5)

p1 <- ggplot(SLBSmod.df, aes(x = .scresid)) +
  geom_histogram(binwidth = .5, color = "white", fill = "gray40") +
  xlab("Standardized Residual") + ylab("Frequency") +
  ggtitle("A") +
  theme_minimal() + theme(text = element_text(size=25))

p2 <- ggplot(SLBSmod.df, aes(x = .fitted, y = .scresid, color = factor(Mo))) +
         geom_point(size = 2) +
        # scale_x_log10() +
         scale_color_discrete(name = "Month") +
  xlab("SLBS [m/s]") + ylab("Standardized Residual") +
  ggtitle("B") +
  theme_minimal() + theme(text = element_text(size=26))
grid.arrange(p1, p2, ncol = 2)
```
Conditional standardized residuals from the linear regression model. (A) Histogram and (B) Residuals as a function of predicted values of SLBS. The pattern in B indicates that predictions for larger values of SLBS are less accurate. 

*How does the model do compared to the actual values? *
```{r}
predictSLBS <- predict(SLBS_m5)
BigDays.sfdfT$predSLBS <- predictSLBS

cor(BigDays.sfdfT$maxBS_shallow, predictSLBS)
```
This model explains 56% of the variation of SLBS.

*Make a table of the model coefficients. *
```{r}
SLBSmod_coefs <- summary(SLBS_m5)
xtable(SLBSmod_coefs$coefficients, digits = 3)
```

*What are the confidence intervals for the model? *
```{r}
confint(SLBS_m5)
```
                 2.5 %      97.5 %
.sig01         2.51850617   5.97403800
.sigma         5.69509099   6.27483241
(Intercept) -312.32914537 -88.86848600
Year           0.04831678   0.15830651
SSTgradient    0.18061139   0.96433173
PNA_10day     -0.92237478  -0.20502250
MJO_10day     -0.40658457   0.06419227

*Plot the observed vs Predicted values of SLBS*
```{r}
ggplot() +
  geom_point(data = BigDays.sfdfT, 
             aes(x = maxBS_shallow, 
                 y = predSLBS, col = Mo), 
             size = 5, 
             show.legend = TRUE,  
             stroke = 0, 
             alpha = 0.8) +
      scale_colour_continuous(type = "viridis") +
  geom_abline(slope = 1, 
                  size = 1.25,
              col = "blue") + 
  ylab("Estimated SLBS [m/s]") + 
      xlab("Observed SLBS [m/s]") +
      theme_minimal() + 
      theme(text = element_text(size=15))
```


#Still working:


##CIN: 
*What is the relationship between these variables and CIN? *
```{r}
CIN_env <- data.frame(AirTemperature = BigDays.sfdfT$avg_globalairtemp, SSTgradient = BigDays.sfdfT$SSTgradient, ElNino = BigDays.sfdfT$avg_Nino3410day, NAO = BigDays.sfdfT$NAO_10day, PNA = BigDays.sfdfT$PNA_10day, MJO = BigDays.sfdfT$MJO_10day, CIN = BigDays.sfdfT$minCIN)


corrplot.mixed(cor(CIN_env), tl.col="black", lower.col="black", number.cex = 1, tl.cex = 1)
```

```{r}
hist(log(BigDays.sfdfT$minCIN*-1))
hist((BigDays.sfdfT$minCIN)**2)
hist(sqrt(BigDays.sfdfT$minCIN*-1))
```

Remove positive values of CIN. Then, get the absolute value of CIN. 
```{r}
sqrtCIN <- sqrt(BigDays.sfdfT$minCIN*-1)
sqrtCIN[is.nan(sqrtCIN)] <- 0
```

```{r}
CIN_m1 <- lmer(sqrtCIN~ Year + avg_globalairtemp + SSTgradient + avg_Nino3410day + NAO_10day + PNA_10day +  MJO_10day + (1|Mo), data = BigDays.sfdfT)
summary(CIN_m1)
ranef(CIN_m1) #There is a strong seasonal variation in CIN values. 
```

```{r}
CIN_m2 <- lmer(sqrtCIN ~ avg_globalairtemp + SSTgradient + avg_Nino3410day + NAO_10day + PNA_10day +  MJO_10day  + (1|Year), data = BigDays.sfdfT)
summary(CIN_m2)
ranef(CIN_m2) #There is a strong annual variation in CIN values  
```

```{r}
CIN_m3 <- glm(sqrtCIN ~ avg_globalairtemp + SSTgradient + avg_Nino3410day + NAO_10day + PNA_10day +  MJO_10day, data = BigDays.sfdfT)
summary(CIN_m3)

```

```{r}
CIN_m4 <- glm(sqrtCIN ~ avg_globalairtemp + SSTgradient + PNA_10day +  MJO_10day, data = BigDays.sfdfT)
summary(CIN_m4)

```

```{r}
CIN_m5 <- glm(sqrtCIN ~ avg_globalairtemp + SSTgradient+ PNA_10day, data = BigDays.sfdfT)
summary(CIN_m5)

```

```{r}
CIN_m6 <- glm(sqrtCIN ~ avg_globalairtemp + SSTgradient, data = BigDays.sfdfT)
summary(CIN_m6)

```
*Compute the AIC*
```{r}
AIC(CIN_m1, CIN_m2, CIN_m3, CIN_m4, CIN_m5, CIN_m6)
```


*Check the residuals of model 6 of CIN.*
```{r}
plot(residuals(CIN_m6)) 
```

*Make a plot of the standardized residuals for the CAPE model*
```{r}
CINmod.df <- BigDays.sfdfT
CINmod.df$predCIN <- predict(CIN_m6)
```

*How does the model do compared to the actual values? *
```{r}
predictCIN <- predict(CIN_m6)
BigDays.sfdfT$predCIN <- predictCIN

cor(BigDays.sfdfT$minCIN, predictCIN*-1)
```
This model explains 18% of the variation of CIN.

*Make a table of the model coefficients. *
```{r}
CINmod_coefs <- summary(CIN_m6)
xtable(CINmod_coefs$coefficients**2, digits = 3)
```

For every one unit increase in the SST gradient, CIN decreases by ~48 J/kg [-85, -10, 95%]. In lay person terms: A larger temperature gradient between Caribbean waters and Gulf of Alaska waters lead to less CIN. 

*What are the confidence intervals for the model? *
```{r}
confint(CIN_m6)
```
                 2.5 %      97.5 %
(Intercept)       -2.3358905 3.9168863
avg_globalairtemp  0.2970740 3.2971301
SSTgradient        0.2777191 0.5860674

*Plot the observed vs Predicted values of CIN*
```{r}
ggplot() +
  geom_point(data = BigDays.sfdfT, 
             aes(x = minCIN, 
                 y = (predCIN**2)*-1, col = Mo), 
             size = 5, 
             show.legend = TRUE,  
             stroke = 0, 
             alpha = 0.8) +
      scale_colour_continuous(type = "viridis") +
  geom_abline(slope = 1, 
                  size = 1.25,
              col = "blue") + 
  ylab("Estimated CIN [J/kg]") + 
      xlab("Observed CIN [J/kg]") +
      theme_minimal() + 
      theme(text = element_text(size=15))
```

*Load the package libraries needed for this research: *
```{r}
suppressMessages(library(dplyr))
suppressMessages(library(sf))
suppressMessages(library(USAboundaries))
suppressMessages(library(tmap))
suppressMessages(library(ggplot2))
suppressMessages(library(lme4))
#devtools::install_github("paul-buerkner/brms")
suppressMessages(library(brms))
#devtools::install_github("rmcelreath/rethinking")
suppressMessages(library(rethinking)) 
suppressMessages(library(tidybayes))
#devtools::install_github("mvuorre/brmstools")
suppressMessages(library(brmstools)) 
suppressMessages(library(bayesplot))
suppressMessages(library(ggpubr))
suppressMessages(library(hexbin))
suppressMessages(library(ggstance))
suppressMessages(library(modelr))
suppressMessages(library(xtable))
suppressMessages(library(sp))
suppressMessages(library(lubridate))
suppressMessages(library(zoo))
```

```{r}
load("BigDays.RData")
```


```{r}
merc =  "+proj=merc +lon_0=0 +k=1 +x_0=0 +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"
```

Create the unique ID for the `All_Tornadoes` file that will match the ID column in `BigDays.sfdfT`. 
```{r}
All_Tornadoes<- All_Tornadoes %>%
   mutate(ID = paste0(gsub("-", "", cDate), groupNumber))
```

##################
## Introduction ##
##################

##########
## Data ##
##########

*Tornado Data*

I load the data. The data file (BigDays.RData) comes from the DefineBigDays repository on github. (https://github.com/zschroder/DefineBigDays). I copy the BigDays.RData file from the DefineBigDays repository and add it to this project. This can be updated yearly by adding the latest tornado csv file from the Storm Prediction Center. Additionally, we can define an outbreak as an arbitrary number of tornadoes.
```{r}
#load("BigDays.RData")
```

Add tornado rate and tornado density to the `BigDays.sfdfT` data frame. 
```{r}
BigDays.sfdfT <- BigDays.sfdfT %>%
  mutate(TorPerHour = nT/as.numeric(Duration) * 3600,
         TorPerKm = nT/as.numeric(HullArea) * 10^6)
dim(BigDays.sfdfT)
```

##########################
## Preliminary Analysis ##
##########################


```{r}
timeseries_env <- BigDays.sfdfT %>%
  group_by(Year) %>%
  summarize(yravgCAPE = mean(maxCAPE),
            yravgCIN = mean(minCIN),
            yravgDLBS = mean(maxBS_deep),
            yravgSLBS = mean(maxBS_shallow))

anom_CAPE = mean(timeseries_env$yravgCAPE)
anom_CIN = mean(timeseries_env$yravgCIN)
anom_DLBS = mean(timeseries_env$yravgDLBS)
anom_SLBS = mean(timeseries_env$yravgSLBS)

timeseries_env <- timeseries_env %>%
  mutate(anomCAPE = yravgCAPE - anom_CAPE,
         anomCIN = yravgCIN - anom_CIN,
         anomDLBS = yravgDLBS - anom_DLBS,
         anomSLBS = yravgSLBS - anom_SLBS,
         posCAPE = anomCAPE >= 0,
         posCIN = anomCIN >= 0,
         posDLBS = anomDLBS >= 0,
         posSLBS = anomSLBS >= 0)
```

## Plot the figures for the yearly average extreme value for CAPE, CIN, DLBS, and SLBS. 

Graph of yearly average maximum CAPE.
```{r, eval = FALSE}
yearCAPEgraph <- ggplot(timeseries_env, (aes(x = Year, y = yravgCAPE))) +
 geom_smooth(method = lm, span = .5, se = FALSE, color = "gray70", size = 1) +
  geom_line(color = "black", lwd = 1) +
  scale_x_continuous(expand = c(0, 0), breaks = c(seq(1994,2018,1))) +
  scale_y_continuous(limits = c(1500, 3000), breaks = c(seq(1500,3000,250))) + 
#  geom_smooth(aes(x = week, y = probs/100), span = .5, se = FALSE, color = "gray70", size = 1) +
  theme_bw() +
  xlab("Year") +
  ylab("Monthly Average of Extreme CAPE (in Clusters)")

yearCAPEgraph <- yearCAPEgraph + theme(axis.title.y = element_text(colour = "black"), axis.text.y = element_text(color = "black"),  axis.text.x = element_text(angle = 45, hjust = 1), axis.text = element_text(size = 12), axis.title=element_text(size=14))

yearCAPEgraph
```

Graph of yearly average minimum CIN.
```{r, eval = FALSE}
yearCINgraph <- ggplot(timeseries_env, (aes(x = Year, y = yravgCIN))) +
 geom_smooth(method = lm, span = .5, se = FALSE, color = "gray70", size = 1) +
  geom_line(color = "black", lwd = 1) +
  scale_x_continuous(expand = c(0, 0), breaks = c(seq(1994,2018,1))) +
  scale_y_continuous(limits = c(-175, -50), breaks = c(seq(-175,-50,25))) + 
#  geom_smooth(aes(x = week, y = probs/100), span = .5, se = FALSE, color = "gray70", size = 1) +
  theme_bw() +
  xlab("Year") +
  ylab("Monthly Average of Extreme CIN (in Clusters)")

yearCINgraph <- yearCINgraph + theme(axis.title.y = element_text(colour = "black"), axis.text.y = element_text(color = "black"),  axis.text.x = element_text(angle = 45, hjust = 1), axis.text = element_text(size = 12), axis.title=element_text(size=14))

yearCINgraph
```

Graph of yearly average maximum DLBS.
```{r, eval = FALSE}
yearDLBSgraph <- ggplot(timeseries_env, (aes(x = Year, y = yravgDLBS))) +
 geom_smooth(method = lm, span = .5, se = FALSE, color = "gray70", size = 1) +
  geom_line(color = "black", lwd = 1) +
  scale_x_continuous(expand = c(0, 0), breaks = c(seq(1994,2018,1))) +
  scale_y_continuous(limits = c(22, 34), breaks = c(seq(22,34,2))) + 
#  geom_smooth(aes(x = week, y = probs/100), span = .5, se = FALSE, color = "gray70", size = 1) +
  theme_bw() +
  xlab("Year") +
  ylab("Monthly Average of Extreme DLBS (in Clusters)")

yearDLBSgraph <- yearDLBSgraph + theme(axis.title.y = element_text(colour = "black"), axis.text.y = element_text(color = "black"),  axis.text.x = element_text(angle = 45, hjust = 1), axis.text = element_text(size = 12), axis.title=element_text(size=14))

yearDLBSgraph
```

Graph of yearly average maximum SLBS.
```{r, eval = FALSE}
yearSLBSgraph <- ggplot(timeseries_env, (aes(x = Year, y = yravgSLBS))) +
 geom_smooth(method = lm, span = .5, se = FALSE, color = "gray70", size = 1) +
  geom_line(color = "black", lwd = 1) +
  scale_x_continuous(expand = c(0, 0), breaks = c(seq(1994,2018,1))) +
  scale_y_continuous(limits = c(10, 20), breaks = c(seq(10,20,2))) + 
#  geom_smooth(aes(x = week, y = probs/100), span = .5, se = FALSE, color = "gray70", size = 1) +
  theme_bw() +
  xlab("Year") +
  ylab("Monthly Average of Extreme SLBS (in Clusters)")

yearSLBSgraph <- yearSLBSgraph + theme(axis.title.y = element_text(colour = "black"), axis.text.y = element_text(color = "black"),  axis.text.x = element_text(angle = 45, hjust = 1), axis.text = element_text(size = 12), axis.title=element_text(size=14))

yearSLBSgraph
```
```{r}
#library(ggpubr)
ggarrange(yearCAPEgraph, yearCINgraph, yearDLBSgraph, yearSLBSgraph)
```

Graph of yearly anomaly of maximum CAPE.
```{r}
yearCAPEgraph <- ggplot(timeseries_env, aes(x = Year, y = anomCAPE)) +
  geom_bar(stat = "identity", aes(fill = posCAPE)) +
  scale_fill_manual(values = c("red2", "blue2"), guide = FALSE) +
  scale_x_continuous(expand = c(0, 0), breaks = c(seq(1994,2018,1))) +
  geom_label(label = paste("Average:", round(anom_CAPE, digits = 2), paste0('J/kg')), x = 1998, y = -600) +
  theme_bw() +
  xlab("Year") +
  ylab("Anomalies of Extreme CAPE (in Clusters)")

yearCAPEgraph <- yearCAPEgraph + theme(axis.title.y = element_text(colour = "black"), axis.text.y = element_text(color = "black"),  axis.text.x = element_text(angle = 45, hjust = 1), axis.text = element_text(size = 12), axis.title=element_text(size=14))
```

Graph of yearly anomaly of minimum CIN.
```{r}
yearCINgraph <- ggplot(timeseries_env, aes(x = Year, y = anomCIN)) +
  geom_bar(stat = "identity", aes(fill = posCIN)) +
  scale_fill_manual(values = c("red2", "blue2"), guide = FALSE) +
  scale_x_continuous(expand = c(0, 0), breaks = c(seq(1994,2018,1))) +
    geom_label(label = paste("Average:", round(anom_CIN, digits = 2), paste0('J/kg')), x = 1998, y = -50) +
  theme_bw() +
  xlab("Year") +
  ylab("Anomalies of Extreme CIN (in Clusters)")

yearCINgraph <- yearCINgraph + theme(axis.title.y = element_text(colour = "black"), axis.text.y = element_text(color = "black"),  axis.text.x = element_text(angle = 45, hjust = 1), axis.text = element_text(size = 12), axis.title=element_text(size=14))
```

Graph of yearly anomaly of maximum DLBS.
```{r}

yearDLBSgraph <- ggplot(timeseries_env, aes(x = Year, y = anomDLBS)) +
  geom_bar(stat = "identity", aes(fill = posDLBS)) +
  scale_fill_manual(values = c("red2", "blue2"), guide = FALSE) +
  scale_x_continuous(expand = c(0, 0), breaks = c(seq(1994,2018,1))) +
  geom_label(label = paste("Average:", round(anom_DLBS, digits = 2), paste0('m/s')), x = 1998, y = 3) +
  theme_bw() +
  xlab("Year") +
  ylab("Anomalies of Extreme DLBS (in Clusters)")

yearDLBSgraph <- yearDLBSgraph + theme(axis.title.y = element_text(colour = "black"), axis.text.y = element_text(color = "black"),  axis.text.x = element_text(angle = 45, hjust = 1), axis.text = element_text(size = 12), axis.title=element_text(size=14))
```

Graph of yearly anomaly of maximum SLBS.
```{r}
yearSLBSgraph <- ggplot(timeseries_env, aes(x = Year, y = anomSLBS)) +
  geom_bar(stat = "identity", aes(fill = posSLBS)) +
  scale_fill_manual(values = c("red2", "blue2"), guide = FALSE) +
  scale_x_continuous(expand = c(0, 0), breaks = c(seq(1994,2018,1))) +
  geom_label(label = paste("Average:", round(anom_SLBS, digits = 2), paste0('m/s')), x = 1998, y = 3.15) +
  theme_bw() +
  xlab("Year") +
  ylab("Anomalies of Extreme SLBS (in Clusters)")

yearSLBSgraph <- yearSLBSgraph + theme(axis.title.y = element_text(colour = "black"), axis.text.y = element_text(color = "black"),  axis.text.x = element_text(angle = 45, hjust = 1), axis.text = element_text(size = 12), axis.title=element_text(size=14))
```

```{r}
#library(ggpubr)
ggarrange(yearCAPEgraph, yearCINgraph, yearDLBSgraph, yearSLBSgraph)
```

#########################
## Climate Change Data ##
#########################


*Environment Data [NARR]*

*Nino 3.4*

*Global Mean Temperature*

*Sea Surface Temperatures*

Surface Temperatures from National Oceanic and Atmospheric Association's (NOAA) Global Surface Temperature (NOAAGlobalTemp) data provided by the NOAA/OAR/ESRL PSL (\url{https://psl.noaa.gov/data/gridded/data.noaaglobaltemp.html}). NOAAGlobalTemp is a combination of land and ocean temperatures combined into a singular dataset. The data includes monthly surface temperatures (both land and water) from 1880 to present. It has a spatial resolution of 5$^\circ$ by 5$^\circ$. 

```{r}
library(raster)
ncpath <- "~/Desktop/Projects/Dissertation_OutbreakClimatology/"
ncname <- "air.mon.anom"  
ncfname <- paste(ncpath, ncname, ".nc", sep="")

tmp_raster <- brick(ncfname, varname="air")
class(tmp_raster)


Month1994to2020 <- seq(1369,1688, 1)

January1994 <- subset(tmp_raster, 1369)
as.data.frame(January1994) #-- > must cut it to the extent of the US

September2020 <- subset(tmp_raster, 1688)
```

```{r, eval = FALSE}
for(i in Month1994to2020){
  print(i)
  #if year = 1994 and month = 1 then you want subset(tmp_raster, 1369)
  CAPE <- raster(rb, layer = 375)
}
```

Sea Surface Temperatures: 

YOu need to switch this to the western Caribbeann. Check ElsnerWiden2014
```{r}
#download.file(url = "https://www.cpc.ncep.noaa.gov/data/indices/sstoi.atl.indices",
#              destfile = "ATL_SST.csv", mode = "wb")  
              #https://psl.noaa.gov/cgi-bin/data/timeseries/timeseries.pl?ntype=1&var=SST&level=2000&lat1=30&lat2=60&lon1=300&lon2=340&iseas=0&mon1=0&mon2=0&iarea=0&typeout=1&Submit=Create+Timeseries
download.file(url = "https://psl.noaa.gov/cgi-bin/data/timeseries/timeseries.pl?ntype=1&var=SST&level=2000&lat1=5&lat2=20&lon1=300&lon2=330+&iseas=0&mon1=0&mon2=0&iarea=0&typeout=1&Submit=Create+Timeseries" ,
              destfile = "Atlantic_SST.csv", mode = "wb") 
```


Nino 3.4 
https://psl.noaa.gov/data/correlation/nina34.data

#Teleconnection Indices 
```{r}
download.file(url = "ftp://ftp.cpc.ncep.noaa.gov/wd52dg/data/indices/tele_index.nh",
              destfile = "tele.csv", mode = "wb")


download.file(url = "https://psl.noaa.gov/gcos_wgsp/Timeseries/Data/nino34.long.data",
              destfile = "Nino3_4.csv", mode = "wb")
```

Arctic sea ice
```{r, eval = FALSE}
#https://nsidc.org/data/NSIDC-0051/versions/1
```



```{r}
#devtools::install_github("wmurphyrd/fiftystater")
library(fiftystater)

sf_fifty <- sf::st_as_sf(fifty_states, coords = c("long", "lat")) %>% 
  group_by(id, piece) %>% 
  summarize(do_union=FALSE) %>%
  st_cast("POLYGON") %>% 
  ungroup()

plot(sf_fifty["id"])

sf_fifty <- sf::st_as_sf(fifty_states, coords = c("long", "lat")) %>% 
  group_by(id, piece) %>% 
  summarize(do_union = FALSE) %>%
  st_cast("POLYGON") %>% 
  ungroup()
```

```{r}
library(ggplot2)
ggplot() +
  theme_minimal() +
  geom_sf(data = sf_fifty)
```

*Monthly figures (highlighting seasonal trends) of climate data*

## Preliminary figures of the climate data

```{r}
ClimatebyMonth <- BigDays.sfdfT %>%
  group_by(Mo) %>%
  summarize(maxtempanom = max(max_globalairtemp),
            SST_max = max(SSTgradient),
            ElNino_max  = max(monthlyElNino34),
            monthlySIE_max = max(monthlySIE))

ClimatebyYear <- BigDays.sfdfT %>%
  group_by(Mo, Year) %>%
  summarize(maxtempanom = max(max_globalairtemp),
            SST_max = max(SSTgradient),
            ElNino_max  = max(monthlyElNino34),
            monthlySIE_max = max(monthlySIE))
```
```{r}
 slopevals = numeric()
slopeargs = numeric()
for (i in 1994:2020) {
      dat <- ClimatebyYear %>% filter(Year == i)
      model = lm(maxtempanom ~ Mo, data = dat)
      slope = model$coefficients[2]
      slopearg = ifelse(model$coefficients[2] >= 0,  "POS", "NEG")
      slopeargs = append(slopeargs, slopearg)
      slopevals = append(slopevals, slope)
}

slope_group = character()
slope_vals = numeric()

for (i in 1:dim(ClimatebyYear)[1]) {
  yr = ClimatebyYear$Year[i]
  yearfordat = (yr - 1994)
  slope_group = append(slope_group, as.data.frame(slopeargs)[1 + yearfordat,])
  slope_vals = append(slope_vals, as.data.frame(slopevals)[1 + yearfordat,])
}

ClimatebyYear$group_Temp <- slope_group
ClimatebyYear$slope_Temp <- slope_vals
```

```{r}
month <- month.abb

yearClimatetempgraph <- ggplot(ClimatebyYear, (aes(x = Mo, y = maxtempanom))) +
 geom_smooth(aes(color=group_Temp), method = lm, span = .5, se = FALSE, size = 1) +
  scale_colour_manual(labels = c("Negative Trend", "Positive Trend"), values = c('red', 'blue')) +
  geom_line(color = "black", lwd = 1) +
  scale_x_continuous(expand = c(0, 0), breaks = c(seq(1,12,1)), labels = month) +
 # scale_y_continuous(limits = c(1500, 3000), breaks = c(seq(1500,3000,250))) + 
#  geom_smooth(aes(x = week, y = probs/100), span = .5, se = FALSE, color = "gray70", size = 1) +
  theme_bw() +
  xlab("Year") +
  ylab("Surface Temperature Anomalies in the U.S.") +
  facet_wrap(~ Year, ncol = 3)

yearClimatetempgraph <- yearClimatetempgraph + theme(axis.title.y = element_text(colour = "black"), axis.text.y = element_text(color = "black"),  axis.text.x = element_text(angle = 60, hjust = 1), axis.text = element_text(size = 10), axis.title=element_text(size=14), legend.title = element_blank(), legend.position = "bottom", legend.key.size = unit(1, "cm"),legend.text=element_text(size=12), panel.spacing = unit(1, "lines"))

yearClimatetempgraph
```


```{r}
 slopevals = numeric()
slopeargs = numeric()
for (i in 1994:2020) {
      dat <- ClimatebyYear %>% filter(Year == i)
      model = lm(SST_max ~ Mo, data = dat)
      slope = model$coefficients[2]
      slopearg = ifelse(model$coefficients[2] >= 0,  "POS", "NEG")
      slopeargs = append(slopeargs, slopearg)
      slopevals = append(slopevals, slope)
}

slope_group = character()
slope_vals = numeric()

for (i in 1:dim(ClimatebyYear)[1]) {
  yr = ClimatebyYear$Year[i]
  yearfordat = (yr - 1994)
  slope_group = append(slope_group, as.data.frame(slopeargs)[1 + yearfordat,])
  slope_vals = append(slope_vals, as.data.frame(slopevals)[1 + yearfordat,])
}

ClimatebyYear$group_SST <- slope_group
ClimatebyYear$slope_SST <- slope_vals
```

```{r}
month <- month.abb

yearSSTgraph <- ggplot(ClimatebyYear, (aes(x = Mo, y = SST_max))) +
 geom_smooth(aes(color=group_SST), method = lm, span = .5, se = FALSE, size = 1) +
  scale_colour_manual(labels = c("Positive Trend", "Negative Trend"), values = c('blue', 'red')) +
  geom_line(color = "black", lwd = 1) +
  scale_x_continuous(expand = c(0, 0), breaks = c(seq(1,12,1)), labels = month) +
 # scale_y_continuous(limits = c(1500, 3000), breaks = c(seq(1500,3000,250))) + 
#  geom_smooth(aes(x = week, y = probs/100), span = .5, se = FALSE, color = "gray70", size = 1) +
  theme_bw() +
  xlab("Year") +
  ylab("Maximum Sea Surface Temperatures in Caribbean [degrees C]") +
  facet_wrap(~ Year, ncol = 3)

yearSSTgraph <- yearSSTgraph + theme(axis.title.y = element_text(colour = "black"), axis.text.y = element_text(color = "black"),  axis.text.x = element_text(angle = 60, hjust = 1), axis.text = element_text(size = 10), axis.title=element_text(size=14), legend.title = element_blank(), legend.position = "bottom", legend.key.size = unit(1, "cm"),legend.text=element_text(size=12), panel.spacing = unit(1, "lines"))

yearSSTgraph
```

```{r}
 slopevals = numeric()
slopeargs = numeric()
for (i in 1994:2020) {
      dat <- ClimatebyYear %>% filter(Year == i)
      model = lm(ElNino_max ~ Mo, data = dat)
      slope = model$coefficients[2]
      slopearg = ifelse(model$coefficients[2] >= 0,  "POS", "NEG")
      slopeargs = append(slopeargs, slopearg)
      slopevals = append(slopevals, slope)
}

slope_group = character()
slope_vals = numeric()

for (i in 1:dim(ClimatebyYear)[1]) {
  yr = ClimatebyYear$Year[i]
  yearfordat = (yr - 1994)
  slope_group = append(slope_group, as.data.frame(slopeargs)[1 + yearfordat,])
  slope_vals = append(slope_vals, as.data.frame(slopevals)[1 + yearfordat,])
}

ClimatebyYear$group_ElNino <- slope_group
ClimatebyYear$slope_ElNino <- slope_vals
```

```{r}
month <- month.abb

yearElNinograph <- ggplot(ClimatebyYear, (aes(x = Mo, y = ElNino_max))) +
 geom_smooth(aes(color=group_ElNino), method = lm, span = .5, se = FALSE, size = 1) +
  scale_colour_manual(labels = c("Negative Trend", "Positive Trend"), values = c('red', 'blue')) +
  geom_line(color = "black", lwd = 1) +
  scale_x_continuous(expand = c(0, 0), breaks = c(seq(1,12,1)), labels = month) +
 # scale_y_continuous(limits = c(1500, 3000), breaks = c(seq(1500,3000,250))) + 
#  geom_smooth(aes(x = week, y = probs/100), span = .5, se = FALSE, color = "gray70", size = 1) +
  theme_bw() +
  xlab("Year") +
  ylab("Maximum El Nino Temperatures [degrees C]") +
  facet_wrap(~ Year, ncol = 3)

yearElNinograph <- yearElNinograph + theme(axis.title.y = element_text(colour = "black"), axis.text.y = element_text(color = "black"),  axis.text.x = element_text(angle = 60, hjust = 1), axis.text = element_text(size = 10), axis.title=element_text(size=14), legend.title = element_blank(), legend.position = "bottom", legend.key.size = unit(1, "cm"),legend.text=element_text(size=12), panel.spacing = unit(1, "lines"))

yearElNinograph
```

```{r}
 slopevals = numeric()
slopeargs = numeric()
for (i in 1994:2020) {
      dat <- ClimatebyYear %>% filter(Year == i)
      model = lm(monthlySIE_max ~ Mo, data = dat)
      slope = model$coefficients[2]
      slopearg = ifelse(model$coefficients[2] >= 0, "POS", "NEG")
      slopeargs = append(slopeargs, slopearg)
      slopevals = append(slopevals, slope)
}

slope_group = character()
slope_vals = numeric()

for (i in 1:dim(ClimatebyYear)[1]) {
  yr = ClimatebyYear$Year[i]
  yearfordat = (yr - 1994)
  slope_group = append(slope_group, as.data.frame(slopeargs)[1 + yearfordat,])
  slope_vals = append(slope_vals, as.data.frame(slopevals)[1 + yearfordat,])
}

ClimatebyYear$group_monthlySIE <- slope_group
ClimatebyYear$slope_monthlySIE <- slope_vals
```

```{r}
month <- month.abb

yearmonthlySIEgraph <- ggplot(ClimatebyYear, (aes(x = Mo, y = monthlySIE_max))) +
 geom_smooth(aes(color=group_monthlySIE), method = lm, span = .5, se = FALSE, size = 1) +
  scale_colour_manual(labels = c("Negative Trend", "Positive Trend"), values = c('red', 'blue')) +
  geom_line(color = "black", lwd = 1) +
  scale_x_continuous(expand = c(0, 0), breaks = c(seq(1,12,1)), labels = month) +
 # scale_y_continuous(limits = c(1500, 3000), breaks = c(seq(1500,3000,250))) + 
#  geom_smooth(aes(x = week, y = probs/100), span = .5, se = FALSE, color = "gray70", size = 1) +
  theme_bw() +
  xlab("Year") +
  ylab("Maximum Arctic Sea Ice Extent [millions of square kilometers]") +
  facet_wrap(~ Year, ncol = 3)

yearmonthlySIEgraph <- yearmonthlySIEgraph + theme(axis.title.y = element_text(colour = "black"), axis.text.y = element_text(color = "black"),  axis.text.x = element_text(angle = 60, hjust = 1), axis.text = element_text(size = 10), axis.title=element_text(size=14), legend.title = element_blank(), legend.position = "bottom", legend.key.size = unit(1, "cm"),legend.text=element_text(size=12), panel.spacing = unit(1, "lines"))

yearmonthlySIEgraph
```