---
title: "Influence of climate indicators on tornado 'outbreak' environments"
author: "Zoe Schroder"
date: "5/1/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

**Make sure to answer these questions in the Introduction: **
Tornadoes will become more destructive in a warmer world. 
    1. How do you know this?
    2. Why is this important?
    3. How does this influence your research?
    
Environmental factors have been linked to individual climate change factors including....
    1. Cite the important literature
    2. How is what you are doing different?
    3. Why is this perspective so important? 

##Introduction: ##
    
Tornado outbreaks are becoming more extensive and destructive on average. Tornado days are declining annually; however, more tornadoes are occurring on these days \citep{ElsnerElsnerJagger2015, TippettEtAl2016, Moore2017, TippettCohen2016, Moore2018, MooreDeBoer2019}. \cite{SchroderElsner2019} use statistical models to quantify the role environmental conditions play in the variability of tornado outbreak characteristics, including accumulated tornado power (ATP), tornado counts, and casualties. Research suggests that ATP and tornado counts is increasing per annum \citep{SchroderElsner2019, ElsnerFrickerSchroder2018}. As a result, the risk for damage increases  \citep{ElsnerMurnaneJaggerWiden2013}. It is challenging to attribute shifts in tornado outbreak characteristics to climate change due to limited understanding between climate change and tornado outbreak environments \citep{ElsnerGuishard2014, ElsnerFrickerSchroder2018}

Researchers project severe convective weather to increase in the future due to more frequent environmental conditions \citep{DelGenioYaoJonas2007, TrappEtAl2007, HoogewindBaldwinTrapp2017, SeeleyRomps2015, DiffenbaughTrappBrooks2008, Brooks2013, SeeleyandRomps2015}. Environmental conditions include a combination of temperature and wind driven factors. CAPE and CIN are temperature-driven factors that are inversely related and represent the ability for air to rise in the atmosphere. CAPE and CIN are projected to increase due to rising surface temperatures \citep{ElsnerJaggerElsner2014, DelGenioYaoJonas2007, HoogewindBaldwinTrapp2017, Brooks2013, TrappEtAl2007, ElsnerFrickerSchroder2018}. BS and SRH are both wind-driven factors. BS represents the change in wind speed and direction over different depths of the atmosphere. BS is projected to decrease due to a limited temperature gradient between the equator and the poles \citep{Brooks2013, DiffenbaughTrappBrooks2008}. SRH represents the rotation within the storm and is projected to increase \citep{TippettEtAl2016}.  Research suggests that the decrease in bulk shear will be less than the projected increase in CAPE \citep{DelGenioYaoJonas2007, TrappEtAl2007, Brooks2013, HoogewindBaldwinTrapp2017}. As a result, favorable severe weather environments are expected to be more frequent \citep{DelGenioYaoJonas2007, TrappEtAl2007, Brooks2013, HoogewindBaldwinTrapp2017}. Researchers suggest shift in environmental factors as a result of climate change, but more work is needed to quantify and explain these potential changes. 

Limited research has been done to link climate change to tornado outbreak environments and associated characteristics. The studies that have been done do not match the scale of the outbreak. Researchers have analyzed climate change's role on severe convective weather at the storm level \citep{TrappandHoogewind2018, HoogewindBaldwinTrapp2017,DelGenioYaoJonas2007, DiffenbaughTrappBrooks2008}. Climate change is a larger scale phenomenon and should be considered with large scale physical processes such as tornado outbreaks. More work is needed to quantify how outbreak-level  environmental conditions are responding to climate change and influencing tornado outbreak characteristics (i.e., cluster severity, tornado counts, and casualty counts). 

The goal of this research is to quantify and explain the relationship between environmental conditions and climate change. We use the term cluster instead of outbreak because we do not associate the tornadoes with the synoptic setup. The objective is to understand the role of climate change on tornado cluster environments and how that change is affecting the climatology of tornado clusters. We will fit statistical models to predict tornado cluster environments using climate change factors, including sea ice extent, El Nino Southern Oscillation Index, global temperature anomalies, and western Caribbean sea surface temperatures. 


```{r}
load("BigDays.RData")
```

**Set the mercator projection for all figures**
```{r}
merc =  "+proj=merc +lon_0=0 +k=1 +x_0=0 +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"
WGS84 <- 4326
```

Load all libraries needed for this research: 
```{r}
library(ggplot2)
library(dplyr)
library(ggpubr)
library(sf)
library(GGally)
library(USAboundaries)
library(tmap)
library(xtable)
library(lme4)
library(gridExtra)
library(corrplot)
```

*Set the coordinate reference system to WGS84*
```{r}
st_crs(BigDays.sfdfT) <- WGS84
```

## Descriptive Statistics

*Big Days: *

```{r}
#The average number of tornadoes per outbreak
mean(BigDays.sfdfT$nT)

#The maximum number of tornadoes per outbreak
max(BigDays.sfdfT$nT)

#What percentage of the clusters have less than the mean number of tornadoes. 
x <- BigDays.sfdfT %>%
  filter(nT <= 22)
dim(x)[1]/dim(BigDays.sfdfT)[1]
```

*What percentage of BigDayTornadoes are associated with each EF rating?*
```{r}
BigTorn = BigDayTornadoes %>%
  group_by(mag) %>%
  summarize(nT = n())

sum(BigDays.sfdfT$n0, na.rm = TRUE)/dim(BigDayTornadoes)[1] *100
#52% of the BigDayTornadoes are EF0
sum(BigDays.sfdfT$n1, na.rm = TRUE)/dim(BigDayTornadoes)[1] *100
#33% of the BigDayTornadoes are EF1
sum(BigDays.sfdfT$n2, na.rm = TRUE)/dim(BigDayTornadoes)[1] *100
#10.5% of the BigDayTornadoes are EF2
sum(BigDays.sfdfT$n3, na.rm = TRUE)/dim(BigDayTornadoes)[1] *100
#3.4% of the BigDayTornadoes are EF3
sum(BigDays.sfdfT$n4, na.rm = TRUE)/dim(BigDayTornadoes)[1] *100
#0.75% of the BigDayTornadoes are EF4
sum(BigDays.sfdfT$n5, na.rm = TRUE)/dim(BigDayTornadoes)[1] *100
#0.08% of the BigDayTornadoes are EF5
```

*what percentage of tornadoes in EF category are contained in a cluster?*
```{r}
AllTorn = All_Tornadoes %>%
  group_by(mag) %>%
  summarize(nT = n())

sum(BigDays.sfdfT$n0, na.rm = TRUE)/AllTorn$nT[1] *100
#48.56% of all EF0s occur in a cluster
sum(BigDays.sfdfT$n1, na.rm = TRUE)/AllTorn$nT[2] *100
#62% of all EF1s occur in a cluster
sum(BigDays.sfdfT$n2, na.rm = TRUE)/AllTorn$nT[3] *100
#73% of all EF2s occur in a cluster
sum(BigDays.sfdfT$n3, na.rm = TRUE)/AllTorn$nT[4] *100
#86% of all EF3s occur in a cluster
sum(BigDays.sfdfT$n4, na.rm = TRUE)/AllTorn$nT[5] *100
#87.5% of all EF4s occur in a cluster
sum(BigDays.sfdfT$n5, na.rm = TRUE)/AllTorn$nT[6] *100
#100% of all EF5s occur in a cluster
```

*Cluster tornadoes*
```{r}
x <- BigDays.sfdfT %>%
  filter(nT==10)
dim(x)[1]
#88 clusters have 10 tornadoes
x <- BigDays.sfdfT %>%
  filter(nT>=30)
dim(x)[1]
#166 clusters have at least 30 tornadoes
x <- BigDays.sfdfT %>%
  filter(nT>=50)
dim(x)[1]
#52 clusters have at least 50 tornadoes. 
```

*Cluster casualties*
```{r}
x <- BigDays.sfdfT %>%
  filter(GroupDayCas==0)
dim(x)[1]
#293 clusters have 0 casualties
x <- BigDays.sfdfT %>%
  filter(GroupDayCas>=30)
dim(x)[1]
#132 clusters have at least 30 casualties
x <- BigDays.sfdfT %>%
  filter(GroupDayCas>=50)
dim(x)[1]
#90 clusters have at least 50 casualties
```

## Preliminary figures of the Environmental Data ##

*Compute the yearly averages and yearly anomalies. *
```{r}
timeseries_env <- as.data.frame(BigDays.sfdfT) %>%
  group_by(Year) %>%
  summarize(yravgCAPE = mean(maxCAPE),
            yravgCIN = mean(minCIN),
            yravgDLBS = mean(maxBS_deep),
            yravgSLBS = mean(maxBS_shallow))

anom_CAPE = mean(timeseries_env$yravgCAPE)
anom_CIN = mean(timeseries_env$yravgCIN)
anom_DLBS = mean(timeseries_env$yravgDLBS)
anom_SLBS = mean(timeseries_env$yravgSLBS)

timeseries_env <- timeseries_env %>%
  mutate(anomCAPE = yravgCAPE - anom_CAPE,
         anomCIN = yravgCIN - anom_CIN,
         anomDLBS = yravgDLBS - anom_DLBS,
         anomSLBS = yravgSLBS - anom_SLBS,
         posCAPE = anomCAPE >= 0,
         posCIN = anomCIN >= 0,
         posDLBS = anomDLBS >= 0,
         posSLBS = anomSLBS >= 0)
```

#Anomaly Figures for CAPE, CIN, DLBS, and SLBS. 

**The Anomalies**

Graph of yearly anomaly of maximum CAPE.
```{r}
yearCAPEgraph <- ggplot(timeseries_env, aes(x = Year, y = anomCAPE)) +
  geom_bar(stat = "identity", aes(fill = posCAPE)) +
  scale_fill_manual(values = c("red2", "blue2"), guide = FALSE) +
  scale_x_continuous(expand = c(0, 0), breaks = c(seq(1994,2020,1))) +
  geom_label(label = paste("Average:", round(anom_CAPE, digits = 2), paste0('J/kg')), x = 1998, y = -600) +
  theme_bw() +
  xlab("Year") +
  ylab("Anomalies of Extreme CAPE (in Clusters)")

yearCAPEgraph <- yearCAPEgraph + theme(axis.title.y = element_text(colour = "black"), axis.text.y = element_text(color = "black"),  axis.text.x = element_text(angle = 45, hjust = 1), axis.text = element_text(size = 12), axis.title=element_text(size=14))
```

Graph of yearly anomaly of minimum CIN.
```{r}
yearCINgraph <- ggplot(timeseries_env, aes(x = Year, y = anomCIN)) +
  geom_bar(stat = "identity", aes(fill = posCIN)) +
  scale_fill_manual(values = c("red2", "blue2"), guide = FALSE) +
  scale_x_continuous(expand = c(0, 0), breaks = c(seq(1994,2020,1))) +
    geom_label(label = paste("Average:", round(anom_CIN, digits = 2), paste0('J/kg')), x = 1998, y = -50) +
  theme_bw() +
  xlab("Year") +
  ylab("Anomalies of Extreme CIN (in Clusters)")

yearCINgraph <- yearCINgraph + theme(axis.title.y = element_text(colour = "black"), axis.text.y = element_text(color = "black"),  axis.text.x = element_text(angle = 45, hjust = 1), axis.text = element_text(size = 12), axis.title=element_text(size=14))
```

Graph of yearly anomaly of maximum DLBS.
```{r}

yearDLBSgraph <- ggplot(timeseries_env, aes(x = Year, y = anomDLBS)) +
  geom_bar(stat = "identity", aes(fill = posDLBS)) +
  scale_fill_manual(values = c("red2", "blue2"), guide = FALSE) +
  scale_x_continuous(expand = c(0, 0), breaks = c(seq(1994,2020,1))) +
  geom_label(label = paste("Average:", round(anom_DLBS, digits = 2), paste0('m/s')), x = 1998, y = 3) +
  theme_bw() +
  xlab("Year") +
  ylab("Anomalies of Extreme DLBS (in Clusters)")

yearDLBSgraph <- yearDLBSgraph + theme(axis.title.y = element_text(colour = "black"), axis.text.y = element_text(color = "black"),  axis.text.x = element_text(angle = 45, hjust = 1), axis.text = element_text(size = 12), axis.title=element_text(size=14))
```

Graph of yearly anomaly of maximum SLBS.
```{r}
yearSLBSgraph <- ggplot(timeseries_env, aes(x = Year, y = anomSLBS)) +
  geom_bar(stat = "identity", aes(fill = posSLBS)) +
  scale_fill_manual(values = c("red2", "blue2"), guide = FALSE) +
  scale_x_continuous(expand = c(0, 0), breaks = c(seq(1994,2020,1))) +
  geom_label(label = paste("Average:", round(anom_SLBS, digits = 2), paste0('m/s')), x = 1998, y = 3.15) +
  theme_bw() +
  xlab("Year") +
  ylab("Anomalies of Extreme SLBS (in Clusters)")

yearSLBSgraph <- yearSLBSgraph + theme(axis.title.y = element_text(colour = "black"), axis.text.y = element_text(color = "black"),  axis.text.x = element_text(angle = 45, hjust = 1), axis.text = element_text(size = 12), axis.title=element_text(size=14))
```

```{r}
#library(ggpubr)
ggarrange(yearCAPEgraph, yearCINgraph, yearDLBSgraph, yearSLBSgraph)
```

#Look at the annual trends in climate variables. 
*Yearly trends in climate variables*
```{r}
library(dplyr)
x <-  BigDays.sfdfT %>%
  group_by(Year) %>%
  summarize(maxmonthlySIE = max(monthlySIE))

monthlySIEgraph <- ggplot(x, (aes(x = Year, y = maxmonthlySIE))) +
 geom_smooth(method = lm, span = .5, se = FALSE, color = "red", size = 1) +
  geom_line(color = "black", lwd = 1) +
  scale_x_continuous(expand = c(0, 0), breaks = c(seq(1994,2020,1))) +
  scale_y_continuous(limits = c(14, 16), breaks = c(seq(14,16,0.25))) + 
#  geom_smooth(aes(x = week, y = probs/100), span = .5, se = FALSE, color = "gray70", size = 1) +
  theme_bw() +
  xlab("Year") +
  ylab("Maximum Monthly Sea Ice Extent [millions of square km]")

monthlySIEgraph <- monthlySIEgraph + theme(axis.title.y = element_text(colour = "black"), axis.text.y = element_text(color = "black"),  axis.text.x = element_text(angle = 45, hjust = 1), axis.text = element_text(size = 12), axis.title=element_text(size=14))
monthlySIEgraph
```

```{r}
x <-  BigDays.sfdfT %>%
  group_by(Year) %>%
  summarize(maxairtemp = max(avg_globalairtemp))

airtempgraph <- ggplot(x, (aes(x = Year, y = maxairtemp))) +
 geom_smooth(method = lm, span = .5, se = FALSE, color = "blue", size = 1) +
  geom_line(color = "black", lwd = 1) +
  scale_x_continuous(expand = c(0, 0), breaks = c(seq(1994,2020,1))) +
  scale_y_continuous(limits = c(0,1), breaks = c(seq(0,1,0.25))) + 
#  geom_smooth(aes(x = week, y = probs/100), span = .5, se = FALSE, color = "gray70", size = 1) +
  theme_bw() +
  xlab("Year") +
  ylab("Monthly Maximum Air Temperature Anomalies")

airtempgraph <- airtempgraph + theme(axis.title.y = element_text(colour = "black"), axis.text.y = element_text(color = "black"),  axis.text.x = element_text(angle = 45, hjust = 1), axis.text = element_text(size = 12), axis.title=element_text(size=14))
airtempgraph
```

```{r}
ggarrange(monthlySIEgraph, airtempgraph, ncol = 2, nrow = 1)
```

```{r}
x <-  BigDays.sfdfT %>%
  group_by(Year) %>%
  summarize(maxCSST = max(avg_CSST10day))

CSSTgraph <- ggplot(x, (aes(x = Year, y = maxCSST))) +
 geom_smooth(method = lm, span = .5, se = FALSE, color = "blue", size = 1) +
  geom_line(color = "black", lwd = 1) +
  scale_x_continuous(expand = c(0, 0), breaks = c(seq(1994,2020,1))) +
  scale_y_continuous(limits = c(28.6,30), breaks = c(seq(28.6,30,.2))) + 
#  geom_smooth(aes(x = week, y = probs/100), span = .5, se = FALSE, color = "gray70", size = 1) +
  theme_bw() +
  xlab("Year") +
  ylab("Maximum 10-day average CSST [degrees C]")

CSSTgraph <- CSSTgraph + theme(axis.title.y = element_text(colour = "black"), axis.text.y = element_text(color = "black"),  axis.text.x = element_text(angle = 45, hjust = 1), axis.text = element_text(size = 12), axis.title=element_text(size=14))
CSSTgraph
```

```{r}
x <-  BigDays.sfdfT %>%
  group_by(Year) %>%
  summarize(maxGAKSST = max(avg_GAKSST10day))

GAKSSTgraph <- ggplot(x, (aes(x = Year, y = maxGAKSST))) +
 geom_smooth(method = lm, span = .5, se = FALSE, color = "blue", size = 1) +
  geom_line(color = "black", lwd = 1) +
  scale_x_continuous(expand = c(0, 0), breaks = c(seq(1994,2020,1))) +
  scale_y_continuous(limits = c(10,15), breaks = c(seq(10,15,.5))) + 
#  geom_smooth(aes(x = week, y = probs/100), span = .5, se = FALSE, color = "gray70", size = 1) +
  theme_bw() +
  xlab("Year") +
  ylab("Maximum 10-day average GAKSST [degrees C]")

GAKSSTgraph <- GAKSSTgraph + theme(axis.title.y = element_text(colour = "black"), axis.text.y = element_text(color = "black"),  axis.text.x = element_text(angle = 45, hjust = 1), axis.text = element_text(size = 12), axis.title=element_text(size=14))
GAKSSTgraph
```

El Nino 10 day average
```{r}
x <-  BigDays.sfdfT %>%
  group_by(Year) %>%
  summarize(maxNino = max(avg_Nino3410day))

Ninograph <- ggplot(x, (aes(x = Year, y = maxNino))) +
 geom_smooth(method = lm, span = .5, se = FALSE, color = "blue", size = 1) +
  geom_line(color = "black", lwd = 1) +
  scale_x_continuous(expand = c(0, 0), breaks = c(seq(1994,2020,1))) +
  scale_y_continuous(limits = c(27, 30), breaks = c(seq(27,30,0.5))) + 
#  geom_smooth(aes(x = week, y = probs/100), span = .5, se = FALSE, color = "gray70", size = 1) +
  theme_bw() +
  xlab("Year") +
  ylab("Maximum 10-day average El Nino 3.4")

Ninograph <- Ninograph + theme(axis.title.y = element_text(colour = "black"), axis.text.y = element_text(color = "black"),  axis.text.x = element_text(angle = 45, hjust = 1), axis.text = element_text(size = 12), axis.title=element_text(size=14))
Ninograph
```

NAO 10 day average
```{r}
x <-  BigDays.sfdfT %>%
  group_by(Year) %>%
  summarize(maxNAO = max(NAO_10day))

NAOgraph <- ggplot(x, (aes(x = Year, y = maxNAO))) +
 geom_smooth(method = lm, span = .5, se = FALSE, color = "blue", size = 1) +
  geom_line(color = "black", lwd = 1) +
  scale_x_continuous(expand = c(0, 0), breaks = c(seq(1994,2020,1))) +
  scale_y_continuous(limits = c(0.5, 3), breaks = c(seq(0.5,3,0.5))) + 
#  geom_smooth(aes(x = week, y = probs/100), span = .5, se = FALSE, color = "gray70", size = 1) +
  theme_bw() +
  xlab("Year") +
  ylab("Maximum 10-day average NAO")

NAOgraph <- NAOgraph + theme(axis.title.y = element_text(colour = "black"), axis.text.y = element_text(color = "black"),  axis.text.x = element_text(angle = 45, hjust = 1), axis.text = element_text(size = 12), axis.title=element_text(size=14))
NAOgraph
```

MJO 10 day average
```{r}
x <-  BigDays.sfdfT %>%
  group_by(Year) %>%
  summarize(maxMJO = max(MJO_10day))

MJOgraph <- ggplot(x, (aes(x = Year, y = maxMJO))) +
 geom_smooth(method = lm, span = .5, se = FALSE, color = "red", size = 1) +
  geom_line(color = "black", lwd = 1) +
  scale_x_continuous(expand = c(0, 0), breaks = c(seq(1994,2020,1))) +
  scale_y_continuous(limits = c(6, 8), breaks = c(seq(6,8,0.5))) + 
#  geom_smooth(aes(x = week, y = probs/100), span = .5, se = FALSE, color = "gray70", size = 1) +
  theme_bw() +
  xlab("Year") +
  ylab(" Maximum 10-day average MJO")

MJOgraph <- MJOgraph + theme(axis.title.y = element_text(colour = "black"), axis.text.y = element_text(color = "black"),  axis.text.x = element_text(angle = 45, hjust = 1), axis.text = element_text(size = 12), axis.title=element_text(size=14))
MJOgraph
```

PNA 10 day average
```{r}
x <-  BigDays.sfdfT %>%
  group_by(Year) %>%
  summarize(maxPNA = max(PNA_10day))

PNAgraph <- ggplot(x, (aes(x = Year, y = maxPNA))) +
 geom_smooth(method = lm, span = .5, se = FALSE, color = "red", size = 1) +
  geom_line(color = "black", lwd = 1) +
  scale_x_continuous(expand = c(0, 0), breaks = c(seq(1994,2020,1))) +
  scale_y_continuous(limits = c(0.5, 4), breaks = c(seq(0.5,4,0.5))) + 
#  geom_smooth(aes(x = week, y = probs/100), span = .5, se = FALSE, color = "gray70", size = 1) +
  theme_bw() +
  xlab("Year") +
  ylab(" Maximum 10-day average PNA")

PNAgraph <- PNAgraph + theme(axis.title.y = element_text(colour = "black"), axis.text.y = element_text(color = "black"),  axis.text.x = element_text(angle = 45, hjust = 1), axis.text = element_text(size = 12), axis.title=element_text(size=14))
PNAgraph
```

```{r}
ggarrange(CSSTgraph, GAKSSTgraph, Ninograph, NAOgraph, MJOgraph, PNAgraph, ncol = 2, nrow = 3)
```

#The Models: 

*Check the distributions of the response variables: *

`CAPE`
```{r}
plot_CAPE <- ggplot(BigDays.sfdfT, aes(maxCAPE)) + 
  geom_histogram(boundary = 50, 
                 bins = 43, 
                 alpha = 0.9, 
                 color = "white", 
                 fill = "Grey40",
                 size = 0.1) +
  scale_x_continuous(expand = c(0,0), limits = c(0,7000)) +
  #scale_x_log10() + 
  xlab("Maximum CAPE [J/kg]") +
  scale_y_continuous(expand = c(0,0), breaks = c(0, 10,20,30,40,50), labels = c("0", "10", "20", "30", "40", "50"), limits = c(0, 50)) +
  ylab("Number of Clusters") + 
  ggtitle("A") +
  theme_bw() +
  theme(text = element_text(size=12), plot.background = element_rect(color = "grey40"), panel.border = element_rect(color = "white"), plot.margin=unit(c(0.5, 0.5, 0.5, 0.5),"cm"))
```

`Deep Layer Shear`
```{r}
plot_DLBS <- ggplot(BigDays.sfdfT, aes(maxBS_deep)) + 
  geom_histogram(boundary = 50, 
                 bins = 43, 
                 alpha = 0.9, 
                 color = "white", 
                 fill = "Grey40",
                 size = 0.1) +
  scale_x_continuous(expand = c(0,0), limits = c(0,50)) +
  #scale_x_log10() + 
  xlab("Maximum Deep Layer Bulk Shear [m/s]") +
  scale_y_continuous(expand = c(0,0), breaks = c(0, 10,20,30,40,50, 60), labels = c("0", "10", "20", "30", "40", "50", "60"), limits = c(0, 60)) +
  ylab("Number of Clusters") + 
  ggtitle("C") +
  theme_bw() +
  theme(text = element_text(size=12), plot.background = element_rect(color = "grey40"), panel.border = element_rect(color = "white"), plot.margin=unit(c(0.5, 0.5, 0.5, 0.5),"cm"))
```

`Shallow Layer Shear`
```{r}
plot_SLBS <- ggplot(BigDays.sfdfT, aes(maxBS_shallow)) + 
  geom_histogram(boundary = 50, 
                 bins = 43, 
                 alpha = 0.9, 
                 color = "white", 
                 fill = "Grey40",
                 size = 0.1) +
  scale_x_continuous(expand = c(0,0), limits = c(0,40)) +
  #scale_x_log10() + 
  xlab("Maximum Shallow Layer Bulk Shear [m/s]") +
  scale_y_continuous(expand = c(0,0), breaks = c(0, 10,20,30,40,50), labels = c("0", "10", "20", "30", "40", "50"), limits = c(0, 50)) +
  ylab("Number of Clusters") + 
  ggtitle("D") +
  theme_bw() +
  theme(text = element_text(size=12), plot.background = element_rect(color = "grey40"), panel.border = element_rect(color = "white"), plot.margin=unit(c(0.5, 0.5, 0.5, 0.5),"cm"))
```

`CIN`
```{r}
plot_CIN <- ggplot(BigDays.sfdfT, aes(minCIN)) + 
  geom_histogram(boundary = 50, 
                 bins = 43, 
                 alpha = 0.9, 
                 color = "white", 
                 fill = "Grey40",
                 size = 0.1) +
  scale_x_continuous(expand = c(0,0), limits = c(-700,50)) +
  #scale_x_log10() + 
  xlab("Minimum CIN [J/kg]") +
  scale_y_continuous(expand = c(0,0), breaks = c(0, 20,40,60,80,100), labels = c("0", "20", "40", "60", "80", "100"), limits = c(0, 100)) +
  ylab("Number of Clusters") + 
  ggtitle("B") +
  theme_bw() +
  theme(text = element_text(size=12), plot.background = element_rect(color = "grey40"), panel.border = element_rect(color = "white"), plot.margin=unit(c(0.5, 0.5, 0.5, 0.5),"cm"))
```

*Combine the distributions into a single figure.*
```{r}
ggarrange(plot_CAPE, plot_CIN, plot_DLBS, plot_SLBS, ncol = 2, nrow = 2)
```

*What is the correlation between the climate variables? *
```{r}
ClimVars <- data.frame( CSST = BigDays.sfdfT$max_CSST10day, GAKSST = BigDays.sfdfT$max_GAKSST10day, ENSO = BigDays.sfdfT$avg_Nino3410day, NAO = BigDays.sfdfT$NAO_10day, PNA = BigDays.sfdfT$PNA_10day, MJO = BigDays.sfdfT$MJO_10day)
```

*Create a figure to show correlations between the climate variables. *
```{r}
corrplot.mixed(cor(ClimVars), tl.col="black", lower.col="black", number.cex = 1, tl.cex = 1)
```

*There is a strong correlation between Caribbean sea surface temperature (ATL_SST) and Gulf of Alaska sea surface temperatures (GAK_SST). Therefore, we need to account for this. We do this by creating a new variable that is the gradient between ATL_SST and GAK_SST. *
```{r,eval=FALSE}
BigDays.sfdfT <- BigDays.sfdfT %>%
  mutate(SSTgradient_10day = max_CSST10day - max_GAKSST10day,
         SSTgradient_20day = max_CSST20day - max_GAKSST20day,
         SSTgradient_30day = max_CSST30day - max_GAKSST30day,
         SSTgradient_5day = max_CSST5day - max_GAKSST5day)
```


What are the values of the climate variables? 
```{r}
max(BigDays.sfdfT$max_Nino3410day)
min(BigDays.sfdfT$max_Nino3410day)
mean(BigDays.sfdfT$max_Nino3410day)

max(BigDays.sfdfT$NAO_10day)
min(BigDays.sfdfT$NAO_10day)
mean(BigDays.sfdfT$NAO_10day)

max(BigDays.sfdfT$PNA_10day)
min(BigDays.sfdfT$PNA_10day)
mean(BigDays.sfdfT$PNA_10day)

max(BigDays.sfdfT$MJO_10day)
min(BigDays.sfdfT$MJO_10day)
mean(BigDays.sfdfT$MJO_10day)

max(BigDays.sfdfT$SSTgradient_10day)
min(BigDays.sfdfT$SSTgradient_10day)
mean(BigDays.sfdfT$SSTgradient_10day)
```

*Now let's look at the correlation matrix again: *
```{r}
ClimVars2 <- data.frame(SSTgradient = BigDays.sfdfT$SSTgradient_10day, ENSO = BigDays.sfdfT$avg_Nino3410day, NAO = BigDays.sfdfT$NAO_10day, PNA = BigDays.sfdfT$PNA_10day, MJO = BigDays.sfdfT$MJO_10day)

corrplot.mixed(cor(ClimVars2), tl.col="black", lower.col="black", number.cex = 1, tl.cex = 1)
```

There is a significant correlation of 0.83 between the SSTgradient and ArcticSIE. `SIE is the monthly Sea Ice Index which provides a quick look at Arctic-wide changes in sea ice. It is a source for consistently processed ice extent and concentration images and data values since 1979.`

This relationship indicates that the stronger the gradient in SST between the Caribbean and the Gulf of Alaska the more Arctic Sea Ice. The large correlation between these two climate variables will cause a collinearity issue in our models. As a result, one of these variables should be removed. I remove the ArcticSIE variable because it makes less physical sense. The SST gradient implies a larger magnitude of Rossby waves that can influence intense convection across the United States. 

```{r}
ClimVars3 <- data.frame(SSTgradient = BigDays.sfdfT$SSTgradient_10day, ENSO = BigDays.sfdfT$avg_Nino3410day, NAO = BigDays.sfdfT$NAO_10day, PNA = BigDays.sfdfT$PNA_10day, MJO = BigDays.sfdfT$MJO_10day)


corrplot.mixed(cor(ClimVars3), tl.col="black", lower.col="black", number.cex = 1, tl.cex = 1)
```

All collinearity issues have been removed. 

##DLBS: 
*What is the relationship between these variables and DLBS? *
```{r}
DLBS_env <- data.frame(SSTgradient = BigDays.sfdfT$SSTgradient_10day, ENSO = BigDays.sfdfT$avg_Nino3410day, NAO = BigDays.sfdfT$NAO_10day, PNA = BigDays.sfdfT$PNA_10day, MJO = BigDays.sfdfT$MJO_10day, DLBS = BigDays.sfdfT$maxBS_deep)


corrplot.mixed(cor(DLBS_env), tl.col="black", lower.col="black", number.cex = 1, tl.cex = 1)
```

```{r}
DLBS_m1 <- lm(maxBS_deep ~ SSTgradient_10day + avg_Nino3410day + NAO_10day + PNA_10day +  MJO_10day + Lat + Lon, data = BigDays.sfdfT)
summary(DLBS_m1)
```

```{r}
DLBSByMonth <- BigDays.sfdfT %>% 
  group_by(Mo) %>%
  summarize(avgDLBS = mean(maxBS_deep)) 
```

`Make a figure of the Mean DLBS from 1994 to 2020 by month`
```{r}
ggplot(DLBSByMonth, aes(x = as.factor(Mo), y = avgDLBS)) +
  geom_bar(stat = "identity", fill = "gray70") +
  scale_x_discrete(labels = month.name) +
  scale_y_continuous(limits = c(0, NA)) +
  ylab("Number of Tornado Groups\nWith At Least 10 Tornadoes") + xlab("") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

( p0 <- ggplot(DLBSByMonth, aes(x = Mo, y = avgDLBS)) +
  geom_bar(stat = "identity", fill = "gray70") +
 scale_x_continuous(breaks = seq(1, 12, 1), labels = month.abb) +
  scale_y_continuous(limits = c(0, 40)) +
  coord_polar(start =0) +
  labs(x = "Month", y = "Averaged Extremes of Deep-Layer Bulk Shear [m/s]") +
  theme_minimal() )
```

```{r}
DLBS_m2 <- lmer(maxBS_deep ~ SSTgradient_10day + avg_Nino3410day + NAO_10day + PNA_10day +  MJO_10day + Lat + Lon +(1|Mo), data = BigDays.sfdfT)
summary(DLBS_m2)
ranef(DLBS_m2) #There is a strong seasonal variation in DLBS values. 
```

```{r}
DLBS_m3 <- lmer(maxBS_deep ~ SSTgradient_10day + NAO_10day + PNA_10day+ Lat + Lon +(1|Mo), data = BigDays.sfdfT)
summary(DLBS_m3)
ranef(DLBS_m3) 
```

```{r}
DLBS_m4 <- lmer(maxBS_deep ~ SSTgradient_20day + avg_Nino3420day + NAO_20day + PNA_20day +  MJO_20day + Lat + Lon +(1|Mo), data = BigDays.sfdfT)
summary(DLBS_m4)
ranef(DLBS_m4)
```

```{r}
DLBS_m5 <- lmer(maxBS_deep ~ SSTgradient_30day + avg_Nino3430day + NAO_30day + PNA_30day +  MJO_30day + Lat + Lon +(1|Mo), data = BigDays.sfdfT)
summary(DLBS_m5)
ranef(DLBS_m5)
```

```{r}
DLBS_m6 <- lmer(maxBS_deep ~ SSTgradient_5day + avg_Nino345day + NAO_5day + PNA_5day +  MJO_5day + Lat + Lon +(1|Mo), data = BigDays.sfdfT)
summary(DLBS_m6)
ranef(DLBS_m6)
```

```{r}
DLBS_m7 <- lmer(maxBS_deep ~ SSTgradient_10day + avg_Nino3430day + NAO_20day + PNA_30day +  MJO_20day + Lat + Lon + (1|Mo), data = BigDays.sfdfT)
summary(DLBS_m7)
ranef(DLBS_m7)
```

```{r}
DLBS_m8 <- lmer(maxBS_deep ~ SSTgradient_10day + NAO_20day + PNA_30day + Lat + Lon + (1|Mo), data = BigDays.sfdfT)
summary(DLBS_m8)
ranef(DLBS_m8)
```

*Model 3 is the best*
```{r}
AIC(DLBS_m1, DLBS_m2, DLBS_m3, DLBS_m4, DLBS_m5, DLBS_m6, DLBS_m7, DLBS_m8)
```

*Check the residuals of model 3 of DLBS.*
```{r}
plot(residuals(DLBS_m8)) 
```
There is no relationship between the residuals.

*Make a plot of the standardized residuals for the DLBS model*
```{r}
DLBSmod.df <- fortify.merMod(model=DLBS_m8, data = getData(DLBS_m8))
DLBSmod.df$predDLBS <- predict(DLBS_m8)

p1 <- ggplot(DLBSmod.df, aes(x = .scresid)) +
  geom_histogram(binwidth = .5, color = "white", fill = "gray40") +
  xlab("Standardized Residual") + ylab("Frequency") +
  ggtitle("A") +
  theme_minimal() + theme(text = element_text(size=25))

p2 <- ggplot(DLBSmod.df, aes(x = .fitted, y = .scresid, color = factor(Mo))) +
         geom_point(size = 2) +
        # scale_x_log10() +
         scale_color_discrete(name = "Month") +
  xlab("DLBS [m/s]") + ylab("Standardized Residual") +
  ggtitle("B") +
  theme_minimal() + theme(text = element_text(size=26))
grid.arrange(p1, p2, ncol = 2)
```
Conditional standardized residuals from the linear regression model. (A) Histogram and (B) Residuals as a function of predicted values of DLBS. The pattern in B indicates that predictions for larger values of DLBS are less accurate. 

*How does the model do compared to the actual values? *
```{r}
predictDLBS <- predict(DLBS_m3)
BigDays.sfdfT$predDLBS <- predictDLBS

cor(BigDays.sfdfT$maxBS_deep, predictDLBS)
```
This model explains 52% of the variation of DLBS.

*Make a table of the model coefficients. *
```{r}
DLBSmod_coefs <- summary(DLBS_m3)
xtable(DLBSmod_coefs$coefficients, digits = 3)
```

*What are the confidence intervals for the model? *
```{r}
confint(DLBS_m3)
```
                       2.5 %      97.5 %
.sig01             1.8313278  5.06219390
.sigma             5.8474468  6.44400817
(Intercept)        4.4455958 20.67462225
SSTgradient_10day  0.3768084  1.22510951
NAO_10day         -0.9717611  0.03026507
PNA_10day         -1.0492033 -0.30952118

*Plot the observed vs Predicted values of DLBS*
```{r}
ggplot() +
  geom_point(data = BigDays.sfdfT, 
             aes(x = maxBS_deep, 
                 y = predDLBS, col = Mo), 
             size = 5, 
             show.legend = TRUE,  
             stroke = 0, 
             alpha = 0.8) +
      scale_colour_continuous(type = "viridis") +
  geom_abline(slope = 1, 
                  size = 1.25,
              col = "blue") + 
  ylab("Estimated DLBS [m/s]") + 
      xlab("Observed DLBS [m/s]") +
      theme_minimal() + 
      theme(text = element_text(size=15))
```

##CAPE: 
*What is the relationship between these variables and CAPE? *
```{r}
CAPE_env <- data.frame(SSTgradient = BigDays.sfdfT$SSTgradient_10day, ENSO = BigDays.sfdfT$avg_Nino3410day, NAO = BigDays.sfdfT$NAO_10day, PNA = BigDays.sfdfT$PNA_10day, MJO = BigDays.sfdfT$MJO_10day, CAPE = BigDays.sfdfT$maxCAPE)


corrplot.mixed(cor(CAPE_env), tl.col="black", lower.col="black", number.cex = 1, tl.cex = 1)
```

```{r}
CAPE_m1 <- lm(maxCAPE ~  SSTgradient_10day + avg_Nino3410day + NAO_10day + PNA_10day +  MJO_10day , data = BigDays.sfdfT)
summary(CAPE_m1)
hist(CAPE_m1$residuals)
library(sm)
sm.density(CAPE_m1$residuals, model = "Normal")
```

```{r}
CAPEByMonth <- BigDays.sfdfT %>% 
  group_by(Mo) %>%
  summarize(avgCAPE = mean(maxCAPE)) 
```

`Make a figure of the total tornadoes from 1994 to 2018 by month`
```{r}
ggplot(CAPEByMonth, aes(x = as.factor(Mo), y = avgCAPE)) +
  geom_bar(stat = "identity", fill = "gray70") +
  scale_x_discrete(labels = month.name) +
  scale_y_continuous(limits = c(0, NA)) +
  ylab("Number of Tornado Groups\nWith At Least 10 Tornadoes") + xlab("") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

( p0 <- ggplot(CAPEByMonth, aes(x = as.factor(Mo), y = avgCAPE)) +
  geom_bar(stat = "identity", fill = "gray70") +
 scale_x_discrete(breaks = seq(1, 12, 1), labels = month.abb) +
  scale_y_continuous(limits = c(0, 4000)) +
  coord_polar(start =0) +
  labs(x = "Month", y = "Averaged Extremes of CAPE [J/kg]") +
  theme_minimal() )
```

```{r}
CAPE_m2 <- lmer(maxCAPE ~ SSTgradient_10day + avg_Nino3410day + NAO_10day + PNA_10day +  MJO_10day + (1|Mo), data = BigDays.sfdfT)
summary(CAPE_m2)
ranef(CAPE_m2) #There is a strong seasonal variation in CAPE values. 
```

```{r}
CAPE_m3 <- lmer(maxCAPE ~ MJO_10day + (1|Mo), data = BigDays.sfdfT)
summary(CAPE_m3)
ranef(CAPE_m3) #There is a strong annual variation in CAPE values  
```


```{r}
CAPE_m4 <- lmer(maxCAPE ~ SSTgradient_20day + avg_Nino3420day + NAO_20day + PNA_20day +  MJO_20day + (1|Mo), data = BigDays.sfdfT)
summary(CAPE_m4)
ranef(CAPE_m4)
```

```{r}
CAPE_m5 <- lmer(maxCAPE ~ SSTgradient_30day + avg_Nino3430day + NAO_30day + PNA_30day +  MJO_30day + (1|Mo), data = BigDays.sfdfT)
summary(CAPE_m5)
ranef(CAPE_m5)
```

```{r}
CAPE_m6 <- lmer(maxCAPE ~ SSTgradient_5day + avg_Nino345day + NAO_5day + PNA_5day +  MJO_5day + (1|Mo), data = BigDays.sfdfT)
summary(CAPE_m6)
ranef(CAPE_m6)
```

Compare the effect size and t values between the 5, 10, 20, and 30 day climate influencers. Select the most important ones. 

```{r}
CAPE_m7 <- lmer(maxCAPE ~ SSTgradient_20day + avg_Nino345day + NAO_5day + PNA_5day +  MJO_10day + (1|Mo), data = BigDays.sfdfT)
summary(CAPE_m7)
ranef(CAPE_m7) 
```

```{r}
CAPE_m8 <- lmer(maxCAPE ~  SSTgradient_20day + PNA_5day +  MJO_10day + (1|Mo), data = BigDays.sfdfT)
summary(CAPE_m8)
ranef(CAPE_m8) #There is a strong annual variation in CAPE values  
```

```{r}
CAPE_m9<- lmer(maxCAPE ~ PNA_5day + MJO_10day + Lat +  Lon + (1|Mo) , data = BigDays.sfdfT)
summary(CAPE_m9)
ranef(CAPE_m9)
```

```{r}
CAPE_m10<- lmer(maxCAPE ~ PNA_5day + MJO_10day +  Lon + (1|Mo) , data = BigDays.sfdfT)
summary(CAPE_m10)
ranef(CAPE_m10)
```

```{r}
CAPE_m11 <- lm(maxCAPE ~ SSTgradient_20day + avg_Nino345day +  MJO_10day, data = BigDays.sfdfT)
summary(CAPE_m11)
```

```{r}
CAPE_m12 <- lm(maxCAPE ~ Month + SSTgradient_20day + avg_Nino345day +  MJO_10day, data = BigDays.sfdfT)
summary(CAPE_m12)
```

```{r}
CAPE_m13 <- lm(maxCAPE ~ Month +  MJO_10day, data = BigDays.sfdfT)
summary(CAPE_m13)
```

```{r}
AIC(CAPE_m1,CAPE_m2,CAPE_m3,CAPE_m4,CAPE_m5,CAPE_m6,CAPE_m7,CAPE_m8, CAPE_m9,CAPE_m10,CAPE_m11, CAPE_m12, CAPE_m13)
```
Using AIC It would have me believe that model 7 is significantly better than all of the others. 

Is there a relationship between SST 20 day and Nino5day or NAO5day? 

```{r}
cor(BigDays.sfdfT$SSTgradient_20day, BigDays.sfdfT$avg_Nino345day)
# 0.08557304
cor(BigDays.sfdfT$SSTgradient_20day, BigDays.sfdfT$NAO_5day)
#0.0008885289
```

*Check the residuals of model 7 of CAPE.*
```{r}
plot(residuals(CAPE_m9)) 
```
There is no relationship between the residuals.

*Make a plot of the standardized residuals for the CAPE model*
```{r}
CAPEmod.df <- fortify.merMod(model=CAPE_m9, data = getData(CAPE_m9), na.rm = TRUE)
CAPEmod.df$predCAPE <- predict(CAPE_m9)

p1 <- ggplot(CAPEmod.df, aes(x = .scresid)) +
  geom_histogram(binwidth = .5, color = "white", fill = "gray40") +
  xlab("Standardized Residual") + ylab("Frequency") +
  ggtitle("A") +
  theme_minimal() + theme(text = element_text(size=25))

p2 <- ggplot(CAPEmod.df, aes(x = .fitted, y = .scresid, color = factor(Mo))) +
         geom_point(size = 2) +
        # scale_x_log10() +
         scale_color_discrete(name = "Month") +
  xlab("CAPE [J/kg]") + ylab("Standardized Residual") +
  ggtitle("B") +
  theme_minimal() + theme(text = element_text(size=26))
grid.arrange(p1, p2, ncol = 2)
```
Conditional standardized residuals from the linear regression model. (A) Histogram and (B) Residuals as a function of predicted values of CAPE. The pattern in B indicates that predictions for larger values of CAPE are less accurate. 

*How does the model do compared to the actual values? *
```{r}
predictCAPE <- predict(CAPE_m9)
BigDays.sfdfT$predCAPE <- predictCAPE

cor(BigDays.sfdfT$maxCAPE, predictCAPE)
```
This model explains 58% of the variation of CAPE.

*Make a table of the model coefficients. *
```{r}
CAPEmod_coefs <- summary(CAPE_m9)
xtable(CAPEmod_coefs$coefficients**2, digits = 3)
```
To read these, you must square them. For every unit increase in the SST gradient, CAPE increases by 24%. (Is that right?)

*What are the confidence intervals for the model? *
```{r}
confint(CAPE_m9)
```
.sig01             542.000800 1271.4251290
.sigma            1059.427473 1167.2740953
(Intercept)       -665.887541 7065.7949915
SSTgradient_20day -126.280940   36.3704786
avg_Nino345day    -141.606635   81.1594236
NAO_5day           -67.578103   95.5077051
PNA_5day          -122.251447    0.3850595
MJO_10day            1.496223   89.6079778

*Plot the observed vs Predicted values of CAPE*
```{r}
ggplot() +
  geom_point(data = BigDays.sfdfT, 
             aes(x = maxCAPE, 
                 y = predCAPE, col = Mo), 
             size = 5, 
             show.legend = TRUE,  
             stroke = 0, 
             alpha = 0.8) +
      scale_colour_continuous(type = "viridis") +
  geom_abline(slope = 1, 
                  size = 1.25,
              col = "blue") + 
  ylab("Estimated CAPE [J/kg]") + 
      xlab("Observed CAPE [J/kg]") +
      theme_minimal() + 
      theme(text = element_text(size=15))
```
