---
title: "Influence of climate indicators on tornado 'outbreak' environments"
author: "Zoe Schroder"
date: "5/1/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

**Make sure to answer these questions in the Introduction: **
Tornadoes will become more destructive in a warmer world. 
    1. How do you know this?
    2. Why is this important?
    3. How does this influence your research?
    
Environmental factors have been linked to individual climate change factors including....
    1. Cite the important literature
    2. How is what you are doing different?
    3. Why is this perspective so important? 

##Introduction: ##
    
Tornado outbreaks are becoming more extensive and destructive on average. Tornado days are declining annually; however, more tornadoes are occurring on these days \citep{ElsnerElsnerJagger2015, TippettEtAl2016, Moore2017, TippettCohen2016, Moore2018, MooreDeBoer2019}. \cite{SchroderElsner2019} use statistical models to quantify the role environmental conditions play in the variability of tornado outbreak characteristics, including accumulated tornado power (ATP), tornado counts, and casualties. Research suggests that ATP and tornado counts is increasing per annum \citep{SchroderElsner2019, ElsnerFrickerSchroder2018}. As a result, the risk for damage increases  \citep{ElsnerMurnaneJaggerWiden2013}. It is challenging to attribute shifts in tornado outbreak characteristics to climate change due to limited understanding between climate change and tornado outbreak environments \citep{ElsnerGuishard2014, ElsnerFrickerSchroder2018}

Researchers project severe convective weather to increase in the future due to more frequent environmental conditions \citep{DelGenioYaoJonas2007, TrappEtAl2007, HoogewindBaldwinTrapp2017, SeeleyRomps2015, DiffenbaughTrappBrooks2008, Brooks2013, SeeleyandRomps2015}. Environmental conditions include a combination of temperature and wind driven factors. CAPE and CIN are temperature-driven factors that are inversely related and represent the ability for air to rise in the atmosphere. CAPE and CIN are projected to increase due to rising surface temperatures \citep{ElsnerJaggerElsner2014, DelGenioYaoJonas2007, HoogewindBaldwinTrapp2017, Brooks2013, TrappEtAl2007, ElsnerFrickerSchroder2018}. BS and SRH are both wind-driven factors. BS represents the change in wind speed and direction over different depths of the atmosphere. BS is projected to decrease due to a limited temperature gradient between the equator and the poles \citep{Brooks2013, DiffenbaughTrappBrooks2008}. SRH represents the rotation within the storm and is projected to increase \citep{TippettEtAl2016}.  Research suggests that the decrease in bulk shear will be less than the projected increase in CAPE \citep{DelGenioYaoJonas2007, TrappEtAl2007, Brooks2013, HoogewindBaldwinTrapp2017}. As a result, favorable severe weather environments are expected to be more frequent \citep{DelGenioYaoJonas2007, TrappEtAl2007, Brooks2013, HoogewindBaldwinTrapp2017}. Researchers suggest shift in environmental factors as a result of climate change, but more work is needed to quantify and explain these potential changes. 

Limited research has been done to link climate change to tornado outbreak environments and associated characteristics. The studies that have been done do not match the scale of the outbreak. Researchers have analyzed climate change's role on severe convective weather at the storm level \citep{TrappandHoogewind2018, HoogewindBaldwinTrapp2017,DelGenioYaoJonas2007, DiffenbaughTrappBrooks2008}. Climate change is a larger scale phenomenon and should be considered with large scale physical processes such as tornado outbreaks. More work is needed to quantify how outbreak-level  environmental conditions are responding to climate change and influencing tornado outbreak characteristics (i.e., cluster severity, tornado counts, and casualty counts). 

The goal of this research is to quantify and explain the relationship between environmental conditions and climate change. We use the term cluster instead of outbreak because we do not associate the tornadoes with the synoptic setup. The objective is to understand the role of climate change on tornado cluster environments and how that change is affecting the climatology of tornado clusters. We will fit statistical models to predict tornado cluster environments using climate change factors, including sea ice extent, El Nino Southern Oscillation Index, global temperature anomalies, and western Caribbean sea surface temperatures. 

```{r}
load("BigDays.RData")
```

**Set the mercator projection for all figures**
```{r}
merc =  "+proj=merc +lon_0=0 +k=1 +x_0=0 +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"
WGS84 <- 4326
```

Load all libraries needed for this research: 
```{r}
library(ggplot2)
library(dplyr)
library(ggpubr)
library(sf)
library(GGally)
library(USAboundaries)
library(tmap)
library(xtable)
library(lme4)
```

*Set the coordinate reference system to WGS84*
```{r}
st_crs(BigDays.sfdfT) <- WGS84
```

#Information on the data: 

#Big Days: 

```{r}
#The average number of tornadoes per outbreak
mean(BigDays.sfdfT$nT)

#The maximum number of tornadoes per outbreak
max(BigDays.sfdfT$nT)

#What percentage of the clusters have less than the mean number of tornadoes. 
x <- BigDays.sfdfT %>%
  filter(nT <= 22)
dim(x)[1]/dim(BigDays.sfdfT)[1]
```
*What percentage of BigDayTornadoes are associated with each EF rating?*
```{r}
BigTorn = BigDayTornadoes %>%
  group_by(mag) %>%
  summarize(nT = n())

sum(BigDays.sfdfT$n0, na.rm = TRUE)/dim(BigDayTornadoes)[1] *100
#52% of the BigDayTornadoes are EF0
sum(BigDays.sfdfT$n1, na.rm = TRUE)/dim(BigDayTornadoes)[1] *100
#33% of the BigDayTornadoes are EF1
sum(BigDays.sfdfT$n2, na.rm = TRUE)/dim(BigDayTornadoes)[1] *100
#10.5% of the BigDayTornadoes are EF2
sum(BigDays.sfdfT$n3, na.rm = TRUE)/dim(BigDayTornadoes)[1] *100
#3.4% of the BigDayTornadoes are EF3
sum(BigDays.sfdfT$n4, na.rm = TRUE)/dim(BigDayTornadoes)[1] *100
#0.75% of the BigDayTornadoes are EF4
sum(BigDays.sfdfT$n5, na.rm = TRUE)/dim(BigDayTornadoes)[1] *100
#0.08% of the BigDayTornadoes are EF5
```

*what percentage of tornadoes in EF category are contained in a cluster?*
```{r}
AllTorn = All_Tornadoes %>%
  group_by(mag) %>%
  summarize(nT = n())

sum(BigDays.sfdfT$n0, na.rm = TRUE)/AllTorn$nT[1] *100
#48.56% of all EF0s occur in a cluster
sum(BigDays.sfdfT$n1, na.rm = TRUE)/AllTorn$nT[2] *100
#62% of all EF1s occur in a cluster
sum(BigDays.sfdfT$n2, na.rm = TRUE)/AllTorn$nT[3] *100
#73% of all EF2s occur in a cluster
sum(BigDays.sfdfT$n3, na.rm = TRUE)/AllTorn$nT[4] *100
#86% of all EF3s occur in a cluster
sum(BigDays.sfdfT$n4, na.rm = TRUE)/AllTorn$nT[5] *100
#87.5% of all EF4s occur in a cluster
sum(BigDays.sfdfT$n5, na.rm = TRUE)/AllTorn$nT[6] *100
#100% of all EF5s occur in a cluster
```


**Extract the geographic centroids of tornado genesis on big tornado days by creating an `sf` object using `st_centroid()` that reduces the MULTIPOINT geometry to a POINT geometry. Compute the area of the group and the number of tornadoes per area (density).**
```{r}
groupDayCentroids.sfdfT <- st_centroid(BigDays.sfdfT)
groupDayCentroids.sfdfT$groupArea <- st_area(st_convex_hull(BigDays.sfdfT))
groupDayCentroids.sfdfT$groupDensity <- groupDayCentroids.sfdfT$nT/groupDayCentroids.sfdfT$groupArea
```

*Get the state borders for figure purposes. Remove Alaska and Hawaii because this research focuses on the contiguous US.*
```{r}
sts <- state.name[!state.name %in% c("Alaska", "Hawaii")]
stateBorders <- us_states(states = sts)
```


*Extract the Hull, centroids, and tornadoes for the May 27, 1996 big day. *
```{r}
May27 <- BigDays.sfdfT %>%
  filter(cDate == "1996-05-27")
May27 <- st_convex_hull(May27)

May27centroid <- groupDayCentroids.sfdfT %>%
  filter(cDate == "1996-05-27")

May27tornadoes <- BigDayTornadoes %>% 
  filter(groupNumber == May27centroid$groupNumber)

May27tornadoes <- May27tornadoes %>%
  mutate(Hour2 = ifelse(Hour <= 6, Hour + 24, Hour))
```

**Make a map of the May 27 tornado day.**
```{r}
May27tornadoes$Hour2 <- cut(as.numeric(May27tornadoes$Hour2), breaks=c(6,12,18,24,30))

May1996 <- (tm_shape(stateBorders) + 
  tm_borders(col = "gray70", alpha = 1) +
  tm_compass(size = 3, fontsize = 1, lwd = 2, color.dark = "gray70") + 
  tm_scale_bar(width = .3, size = 0.8, lwd = 1.75, color.dark = "gray70") +
  tm_layout(legend.bg.color = "white", 
            legend.text.size = .75, 
            attr.position = c("left", "bottom"), 
            inner.margins = c(.2, .2, .2, .2)) +
tm_shape(May27, is.master = TRUE, projection = "+proj=merc +lon_0=0 +k=1 +x_0=0 +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs") +
  tm_borders(col = "black", lwd = 3) +
tm_shape(May27tornadoes) +
  tm_symbols(size = 4, col = "Hour2", alpha = 0.8, palette = "BuPu", title.col = "Time [CST]", labels = c("6 to 12", "12 to 18", "18 to 24", "0 to 6"), border.alpha = 0) +
    tm_layout(legend.title.size = 1.1,
            legend.position = c("right", "bottom"), 
            legend.stack = "horizontal",
            legend.frame = FALSE, 
            legend.text.size = 1, legend.width = -0.2) +
tm_shape(May27centroid) +
  tm_symbols(size = 1.25, col = "black", shape = 24) +
  tm_layout(title = "May 27, 1996 \n 10 tornadoes", 
              title.position = c("center", "top"), 
              legend.title.size = 1.4,
              legend.position = c("right", "bottom"), 
              legend.stack = "horizontal",
              legend.frame = FALSE, 
              legend.text.size = 1.2, 
              legend.width = -0.2, 
              title.size = 1.5))
May1996
```
*Extract the Hull, centroids, and tornadoes for the May 5, 2007 big day. *
```{r}
May05 <- BigDays.sfdfT %>%
  filter(cDate == "2007-05-05")
May05 <- st_convex_hull(May05)

May05centroid <- groupDayCentroids.sfdfT %>%
  filter(cDate == "2007-05-05")

May05tornadoes <- BigDayTornadoes %>% 
  filter(groupNumber == May05centroid$groupNumber)

May05tornadoes <- May05tornadoes %>%
  mutate(Hour2 = ifelse(Hour <= 6, Hour + 24, Hour))
```

**Make a map of the May 5 tornado day.**
```{r}
May05tornadoes$Hour2 <- cut(May05tornadoes$Hour2, breaks=c(6,12,18,24,30))

May2007 <- (tm_shape(stateBorders) + 
  tm_borders(col = "gray70", alpha = 1) +
  tm_compass(size = 3, fontsize = 1, lwd = 2, color.dark = "gray70") +       
  tm_scale_bar(width = .3, size = 0.8, lwd = 1.75, color.dark = "gray70") +
  tm_layout(legend.bg.color = "white", 
            legend.text.size = .75, 
            attr.position = c("left", "bottom"), 
            inner.margins = c(.15, .15, .15, .15)) +
tm_shape(May05, is.master = TRUE, projection = "+proj=merc +lon_0=0 +k=1 +x_0=0 +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs") +
  tm_borders(col = "black", lwd = 3) +
tm_shape(May05tornadoes) +
  tm_symbols(size = 4, col = "Hour2", alpha = 0.8, palette = "BuPu", title.col = "Time [CST]", labels = c("6 to 12", "12 to 18", "18 to 24", "0 to 6"), border.alpha = 0) +
    tm_layout(legend.title.size = 1.1,
            legend.position = c("right", "bottom"), 
            legend.stack = "horizontal",
            legend.frame = FALSE, 
            legend.text.size = 1, legend.width = -0.2) +
tm_shape(May05centroid) +
  tm_symbols(size = 1.25, col = "black", shape = 24)  +
  tm_layout(title = "May 5, 2007 \n 90 tornadoes", 
              title.position = c("center", "top"), 
              legend.title.size = 1.4,
              legend.position = c("right", "bottom"), 
              legend.stack = "horizontal",
              legend.frame = FALSE, 
              legend.text.size = 1.2, 
              legend.width = -0.2, 
              title.size = 1.5)
)

May2007
```
*Combine the two cluster examples into a single figure.*
```{r}
tmap_arrange(May1996, May2007)
```

Environmental Data: 

*Cluster statistics by time of day. Each cluster is categorized by the closest three-hour time (defined by the NARR data) prior to the first tornado.*
```{r}
BigDays.sfdfT %>%
  group_by(NARRZtime) %>%
  summarize(NumClus = n(),
            NumTorns = sum(nT),
            AvgTornsperClus = NumTorns/NumClus,
            AvgDur = mean(Duration)/3600)

#How many clusters occur between 18 and 0 Z?
(258 + 218)/830
```

*May 6, 2003 Big Day*
```{r}
May6 <- BigDays.sfdfT %>%
  filter(ID == "200305062624")
May6th <- st_convex_hull(May6)

May6centroid <- groupDayCentroids.sfdfT %>%
  filter(ID == "200305062624")

May6tornadoes <- All_Tornadoes %>% 
  filter(ID == "20030506" & groupNumber == "2624")
```

```{r, eval = FALSE}
tm_shape(stateBorders) + 
  tm_borders(col = "gray70", alpha = 1) +
  tm_scale_bar(width = .3, size = 0.8, lwd = 1.75, color.dark = "gray70") +
  tm_compass(size = 3, fontsize = 1, lwd = 2, color.dark = "gray70") + 
  tm_layout(legend.bg.color = "white", 
            legend.text.size = .75, 
            attr.position = c("right", "top"), 
            inner.margins = c(.15, .15, .1, .2)) +
#tm_shape(counties.sf) +
#  tm_borders(col = "gray40", alpha = .3) +
#  tm_scale_bar(width = 8, size = 8, color.dark = "gray70") +
  #tm_format("World", legend.position = c("right", "top"),
#                   attr.position = c("right", "top"),
#                   legend.frame = FALSE,
                   #title = "May 30th Tornado Group",
                   #title.size = 1.3,
                   #title.position = c("left", "TOP"),
 #                  inner.margins = c(.2, .2, .2, .2)) +
tm_shape(May6, is.master = TRUE, projection = merc) +
  tm_borders(col = "black", lwd = 3) +
tm_shape(May6tornadoes) +
  tm_symbols(size = 2, col = "purple4", alpha = 0.8, border.alpha = 0) +
    tm_layout(legend.title.size = 1.1,
            legend.position = c("right", "bottom"), 
            legend.stack = "horizontal",
            legend.frame = FALSE, 
            legend.text.size = 1, legend.width = -0.2) +
tm_shape(May6centroid) +
  tm_symbols(size = 1.25, col = "black", shape = 24) 
```

*Plot the CAPE, HLCY, and CIN for the May 6, 2003 Big Day. First, read in the raster. *
```{r, eval = FALSE}
library(raster)
 rb <- brick(paste0("/Users/zoeschroder/Desktop/Projects/Dissertation_OutbreakClimatology/merged_AWIP32.20030506/merged_AWIP32.2003050612")) #<-- this is for varying NARR times
  CAPE <- raster(rb, layer = 375)
  HLCY <- raster(rb, layer = 323)
  CIN <- raster(rb, layer = 376)
  UGRD500 <- raster(rb, layer = 117) 
  VGRD500 <- raster(rb, layer = 118) 
  UGRDsfc <- raster(rb, layer = 293) 
  VGRDsfc <- raster(rb, layer = 294)     
  BS <- sqrt(((UGRD500 - UGRDsfc)**2) + ((VGRD500 - VGRDsfc)**2))
```
**Convective Available Potential energy **
```{r, eval = FALSE}
CAPE_May6 <- crop(CAPE, extent(May6))
CAPE_extent<- mask(CAPE_May6, May6)
plot(CAPE_extent)

maxCAPE <- which.max(CAPE_extent) #Returns a pixel number

#Test that this max value equals the one in BigDays.sfdfT
test <- getValues(CAPE_extent) 
test<-as.matrix(test, ncol=1)
```

```{r, eval = FALSE}
CAPEmaxpos <- xyFromCell(CAPE_extent, maxCAPE)

xy <- data.frame(CAPEmaxpos)
CAPE_latlon <- data.frame(lon=xy$x, lat=xy$y)
print(CAPE_latlon)

CAPE_latlon <- SpatialPoints(CAPE_latlon)
proj4string(CAPE_latlon) = CRS("+proj=lcc +lat_1=50 +lat_2=50 +lat_0=50 +lon_0=-107 +x_0=0 +y_0=0 +a=6371200 +b=6371200 +units=m +no_defs")
```

```{r, eval = FALSE}
counties <- us_counties()
states <- us_states()
```

```{r, eval = FALSE}
p1 <- tm_shape(CAPE_extent) +
  tm_raster(title = "",  n = 6, palette = "BuPu") +
  tm_format(title = expression(paste("CAPE [J ", kg^-1, "]")), "World", legend.position = c("right", "bottom"),
                   attr.position = c("right", "bottom"),
                   legend.frame = FALSE,
                   inner.margins = c(.15, 0, .15, .18), #bottom, left, top
                    legend.text.size = 1, legend.width = -0.28, legend.title.size=1, legend.bg.color = "white", title.size = 1.4) +
#tm_shape(counties) +
  #tm_borders(col = "grey") +
  tm_scale_bar(width = 0.3, size = 0.8, position = c("right", "top"), color.dark = "gray70") +
tm_shape(states) +
  tm_borders() +
tm_shape(May6, is.master = TRUE, projection = merc) +
  tm_borders(col = "black", lwd = 3)  +
#tm_shape(May6centroid) +
#  tm_symbols(size = 1.25, col = "black", shape = 24) +
tm_shape(CAPE_latlon) + 
  tm_symbols(size = 1.4, col = "black", shape = 22)
p1
```

**Helicity**
```{r, eval = FALSE}
HLCY_May6 <- crop(HLCY, extent(May6))
HLCY_extent<- mask(HLCY_May6, May6)
plot(HLCY_extent)

maxHLCY <- which.max(HLCY_extent)
```

```{r, eval = FALSE}
HLCYmaxpos <- xyFromCell(HLCY_extent, maxHLCY)

xy <- data.frame(HLCYmaxpos)
HLCY_latlon <- data.frame(lon=xy$x, lat=xy$y)
print(HLCY_latlon)

HLCY_latlon <- SpatialPoints(HLCY_latlon)
proj4string(HLCY_latlon) = CRS("+proj=lcc +lat_1=50 +lat_2=50 +lat_0=50 +lon_0=-107 +x_0=0 +y_0=0 +a=6371200 +b=6371200 +units=m +no_defs")
```

```{r, eval = FALSE}
p2 <- tm_shape(HLCY_extent) +
  tm_raster(title = "", n = 5, palette = "BuPu") + #"YlGnBu") +
  tm_format(title = expression(paste("Helicity [",m^2, s^-2, "]")), "World", legend.position = c("right", "bottom"),
                   attr.position = c("right", "bottom"),
                   legend.frame = FALSE,
                   inner.margins = c(.15, 0, .15, .15), #bottom, left, top
                    legend.text.size = 1, legend.width = -0.25, legend.title.size=1, legend.bg.color = "white", title.size = 1.4) +
#tm_shape(counties) +
#  tm_borders(col = "grey") +
  tm_scale_bar(width = 0.3, size = 0.8, position = c("right", "top"), color.dark = "gray70") +
tm_shape(states) +
  tm_borders() +
tm_shape(May6, is.master = TRUE, projection = merc) +
  tm_borders(col = "black", lwd = 3)  +
#tm_shape(May6centroid) +
#  tm_symbols(size = 1.25, col = "black", shape = 24) +
tm_shape(HLCY_latlon) + 
  tm_symbols(size = 1.4, col = "black", shape = 22)
p2
```

**Convective Inhibition **
```{r, eval = FALSE}
CIN_May6 <- crop(CIN, extent(May6))
CIN_extent<- mask(CIN_May6, May6)

minCIN <- which.min(CIN_extent)
```

```{r, eval = FALSE}
CINminpos <- xyFromCell(CIN_extent, minCIN)

xy <- data.frame(CINminpos)
CIN_latlon <- data.frame(lon=xy$x, lat=xy$y)
print(CIN_latlon)

CIN_latlon <- SpatialPoints(CIN_latlon)
proj4string(CIN_latlon) = CRS("+proj=lcc +lat_1=50 +lat_2=50 +lat_0=50 +lon_0=-107 +x_0=0 +y_0=0 +a=6371200 +b=6371200 +units=m +no_defs")
```

```{r, eval = FALSE}
p3 <- tm_shape(CIN_extent) +
  tm_raster(title = "", n =5, palette = "BuPu") + #"YlOrBr") +
  tm_format(title = expression(paste("CIN [J ", kg^-1, "]")), "World", legend.position = c("right", "bottom"),
                   attr.position = c("right", "bottom"),
                   legend.frame = FALSE,
                   inner.margins = c(.15, 0, .15, .15), #bottom, left, top
                    legend.text.size = 1., legend.width = -0.3, legend.title.size=1, legend.bg.color = "white", title.size = 1.4) +
#tm_shape(counties) +
#  tm_borders(col = "grey") +
  tm_scale_bar(width = 0.3, size = 0.8, position = c("right", "top"), color.dark = "gray70") +
tm_shape(states) +
  tm_borders() +
tm_shape(May6, is.master = TRUE, projection = merc) +
  tm_borders(col = "black", lwd = 3)  +
#tm_shape(May6centroid) +
#  tm_symbols(size = 1.25, col = "black", shape = 24) +
tm_shape(CIN_latlon) + 
  tm_symbols(size = 1.4, col = "black", shape = 22)
p3
```

**Bulk Shear**
```{r, eval = FALSE}
BS_May6 <- crop(BS, extent(May6))
BS_extent<- mask(BS_May6, May6)
plot(BS_extent)

maxBS <- which.max(BS_extent)
```

```{r, eval = FALSE}
BSmaxpos <- xyFromCell(BS_extent, maxBS)

xy <- data.frame(BSmaxpos)
BS_latlon <- data.frame(lon=xy$x, lat=xy$y)
print(BS_latlon)

BS_latlon <- SpatialPoints(BS_latlon)
proj4string(BS_latlon) = CRS("+proj=lcc +lat_1=50 +lat_2=50 +lat_0=50 +lon_0=-107 +x_0=0 +y_0=0 +a=6371200 +b=6371200 +units=m +no_defs")
```

```{r, eval = FALSE}
p4 <- tm_shape(BS_extent) +
  tm_raster(title = "", n = 5, palette = "BuPu") +
  tm_format(title = expression(paste("Bulk Shear [m ", s^-1, "]")), "World", legend.position = c("right", "bottom"),
                   attr.position = c("right", "bottom"),
                   legend.frame = FALSE,
                   inner.margins = c(.15, 0, .15, .15), #bottom, left, top
                    legend.text.size = 1, legend.width = -0.2, legend.title.size=1, legend.bg.color = "white", title.size = 1.4) +
#tm_shape(counties) +
#  tm_borders(col = "grey") +
  tm_scale_bar(width = 0.3, size = 0.8, position = c("right", "top"), color.dark = "gray70") +
tm_shape(states) +
  tm_borders() +
tm_shape(May6, is.master = TRUE, projection = merc) +
  tm_borders(col = "black", lwd = 3)  +
#tm_shape(May6centroid) +
#  tm_symbols(size = 1.25, col = "black", shape = 24) +
tm_shape(BS_latlon) + 
  tm_symbols(size = 1.4, col = "black", shape = 22)
p4
```

**Combine into one figure: **
```{r, eval = FALSE}
tmap_arrange(p1, p2, p3, p4, ncol = 2, nrow = 2)
```
`Figure: Environmental conditions at 12 UTC on May 6, 2003. The black line is the spatial extent of the tornado genesis locations and the black triangle is the centroid of the locations. The first tornado in the cluster started at 14:20 UTC. The black square indicates the locations of the highest value of CAPE (3660~J~kg$^{-1}$), the lowest value of CIN ($-$149~J~kg$^{-1}$), the highest value of helicity (308~m$^2$~s$^{-2}$) and the highest value of bulk shear (33~m~s$^{-1}$).`

## Preliminary figures of the Environmental Data ##

*Compute the yearly averages and yearly anomalies. *
```{r}
timeseries_env <- as.data.frame(BigDays.sfdfT) %>%
  group_by(Year) %>%
  summarize(yravgCAPE = mean(maxCAPE),
            yravgCIN = mean(minCIN),
            yravgDLBS = mean(maxBS_deep),
            yravgSLBS = mean(maxBS_shallow))

anom_CAPE = mean(timeseries_env$yravgCAPE)
anom_CIN = mean(timeseries_env$yravgCIN)
anom_DLBS = mean(timeseries_env$yravgDLBS)
anom_SLBS = mean(timeseries_env$yravgSLBS)

timeseries_env <- timeseries_env %>%
  mutate(anomCAPE = yravgCAPE - anom_CAPE,
         anomCIN = yravgCIN - anom_CIN,
         anomDLBS = yravgDLBS - anom_DLBS,
         anomSLBS = yravgSLBS - anom_SLBS,
         posCAPE = anomCAPE >= 0,
         posCIN = anomCIN >= 0,
         posDLBS = anomDLBS >= 0,
         posSLBS = anomSLBS >= 0)
```

#Anomaly Figures for CAPE, CIN, DLBS, and SLBS. 

**The Anomalies**

Graph of yearly anomaly of maximum CAPE.
```{r}
yearCAPEgraph <- ggplot(timeseries_env, aes(x = Year, y = anomCAPE)) +
  geom_bar(stat = "identity", aes(fill = posCAPE)) +
  scale_fill_manual(values = c("red2", "blue2"), guide = FALSE) +
  scale_x_continuous(expand = c(0, 0), breaks = c(seq(1994,2020,1))) +
  geom_label(label = paste("Average:", round(anom_CAPE, digits = 2), paste0('J/kg')), x = 1998, y = -600) +
  theme_bw() +
  xlab("Year") +
  ylab("Anomalies of Extreme CAPE (in Clusters)")

yearCAPEgraph <- yearCAPEgraph + theme(axis.title.y = element_text(colour = "black"), axis.text.y = element_text(color = "black"),  axis.text.x = element_text(angle = 45, hjust = 1), axis.text = element_text(size = 12), axis.title=element_text(size=14))
```

Graph of yearly anomaly of minimum CIN.
```{r}
yearCINgraph <- ggplot(timeseries_env, aes(x = Year, y = anomCIN)) +
  geom_bar(stat = "identity", aes(fill = posCIN)) +
  scale_fill_manual(values = c("red2", "blue2"), guide = FALSE) +
  scale_x_continuous(expand = c(0, 0), breaks = c(seq(1994,2020,1))) +
    geom_label(label = paste("Average:", round(anom_CIN, digits = 2), paste0('J/kg')), x = 1998, y = -50) +
  theme_bw() +
  xlab("Year") +
  ylab("Anomalies of Extreme CIN (in Clusters)")

yearCINgraph <- yearCINgraph + theme(axis.title.y = element_text(colour = "black"), axis.text.y = element_text(color = "black"),  axis.text.x = element_text(angle = 45, hjust = 1), axis.text = element_text(size = 12), axis.title=element_text(size=14))
```

Graph of yearly anomaly of maximum DLBS.
```{r}

yearDLBSgraph <- ggplot(timeseries_env, aes(x = Year, y = anomDLBS)) +
  geom_bar(stat = "identity", aes(fill = posDLBS)) +
  scale_fill_manual(values = c("red2", "blue2"), guide = FALSE) +
  scale_x_continuous(expand = c(0, 0), breaks = c(seq(1994,2020,1))) +
  geom_label(label = paste("Average:", round(anom_DLBS, digits = 2), paste0('m/s')), x = 1998, y = 3) +
  theme_bw() +
  xlab("Year") +
  ylab("Anomalies of Extreme DLBS (in Clusters)")

yearDLBSgraph <- yearDLBSgraph + theme(axis.title.y = element_text(colour = "black"), axis.text.y = element_text(color = "black"),  axis.text.x = element_text(angle = 45, hjust = 1), axis.text = element_text(size = 12), axis.title=element_text(size=14))
```

Graph of yearly anomaly of maximum SLBS.
```{r}
yearSLBSgraph <- ggplot(timeseries_env, aes(x = Year, y = anomSLBS)) +
  geom_bar(stat = "identity", aes(fill = posSLBS)) +
  scale_fill_manual(values = c("red2", "blue2"), guide = FALSE) +
  scale_x_continuous(expand = c(0, 0), breaks = c(seq(1994,2020,1))) +
  geom_label(label = paste("Average:", round(anom_SLBS, digits = 2), paste0('m/s')), x = 1998, y = 3.15) +
  theme_bw() +
  xlab("Year") +
  ylab("Anomalies of Extreme SLBS (in Clusters)")

yearSLBSgraph <- yearSLBSgraph + theme(axis.title.y = element_text(colour = "black"), axis.text.y = element_text(color = "black"),  axis.text.x = element_text(angle = 45, hjust = 1), axis.text = element_text(size = 12), axis.title=element_text(size=14))
```

```{r}
#library(ggpubr)
ggarrange(yearCAPEgraph, yearCINgraph, yearDLBSgraph, yearSLBSgraph)
```

## Preliminary figures of the climate data

```{r}
ClimatebyMonth <- BigDays.sfdfT %>%
  group_by(Mo) %>%
  summarize(maxtempanom = max(max_airtemp),
            SST_max = max(SSTgradient),
            ElNino_max  = max(monthlyElNino34),
            monthlySIE_max = max(monthlySIE))

ClimatebyYear <- BigDays.sfdfT %>%
  group_by(Mo, Year) %>%
  summarize(maxtempanom = max(max_airtemp),
            SST_max = max(SSTgradient),
            ElNino_max  = max(monthlyElNino34),
            monthlySIE_max = max(monthlySIE))
```

*Monthly figures (highlighting seasonal trends) of climate data*
```{r}
 slopevals = numeric()
slopeargs = numeric()
for (i in 1994:2020) {
      dat <- ClimatebyYear %>% filter(Year == i)
      model = lm(maxtempanom ~ Mo, data = dat)
      slope = model$coefficients[2]
      slopearg = ifelse(model$coefficients[2] >= 0,  "POS", "NEG")
      slopeargs = append(slopeargs, slopearg)
      slopevals = append(slopevals, slope)
}

slope_group = character()
slope_vals = numeric()

for (i in 1:dim(ClimatebyYear)[1]) {
  yr = ClimatebyYear$Year[i]
  yearfordat = (yr - 1994)
  slope_group = append(slope_group, as.data.frame(slopeargs)[1 + yearfordat,])
  slope_vals = append(slope_vals, as.data.frame(slopevals)[1 + yearfordat,])
}

ClimatebyYear$group_Temp <- slope_group
ClimatebyYear$slope_Temp <- slope_vals
```

```{r}
month <- month.abb

yearClimatetempgraph <- ggplot(ClimatebyYear, (aes(x = Mo, y = maxtempanom))) +
 geom_smooth(aes(color=group_Temp), method = lm, span = .5, se = FALSE, size = 1) +
  scale_colour_manual(labels = c("Negative Trend", "Positive Trend"), values = c('red', 'blue')) +
  geom_line(color = "black", lwd = 1) +
  scale_x_continuous(expand = c(0, 0), breaks = c(seq(1,12,1)), labels = month) +
 # scale_y_continuous(limits = c(1500, 3000), breaks = c(seq(1500,3000,250))) + 
#  geom_smooth(aes(x = week, y = probs/100), span = .5, se = FALSE, color = "gray70", size = 1) +
  theme_bw() +
  xlab("Year") +
  ylab("Surface Temperature Anomalies in the U.S.") +
  facet_wrap(~ Year, ncol = 3)

yearClimatetempgraph <- yearClimatetempgraph + theme(axis.title.y = element_text(colour = "black"), axis.text.y = element_text(color = "black"),  axis.text.x = element_text(angle = 60, hjust = 1), axis.text = element_text(size = 10), axis.title=element_text(size=14), legend.title = element_blank(), legend.position = "bottom", legend.key.size = unit(1, "cm"),legend.text=element_text(size=12), panel.spacing = unit(1, "lines"))

yearClimatetempgraph
```


```{r}
 slopevals = numeric()
slopeargs = numeric()
for (i in 1994:2020) {
      dat <- ClimatebyYear %>% filter(Year == i)
      model = lm(SST_max ~ Mo, data = dat)
      slope = model$coefficients[2]
      slopearg = ifelse(model$coefficients[2] >= 0,  "POS", "NEG")
      slopeargs = append(slopeargs, slopearg)
      slopevals = append(slopevals, slope)
}

slope_group = character()
slope_vals = numeric()

for (i in 1:dim(ClimatebyYear)[1]) {
  yr = ClimatebyYear$Year[i]
  yearfordat = (yr - 1994)
  slope_group = append(slope_group, as.data.frame(slopeargs)[1 + yearfordat,])
  slope_vals = append(slope_vals, as.data.frame(slopevals)[1 + yearfordat,])
}

ClimatebyYear$group_SST <- slope_group
ClimatebyYear$slope_SST <- slope_vals
```

```{r}
month <- month.abb

yearSSTgraph <- ggplot(ClimatebyYear, (aes(x = Mo, y = SST_max))) +
 geom_smooth(aes(color=group_SST), method = lm, span = .5, se = FALSE, size = 1) +
  scale_colour_manual(labels = c("Positive Trend", "Negative Trend"), values = c('blue', 'red')) +
  geom_line(color = "black", lwd = 1) +
  scale_x_continuous(expand = c(0, 0), breaks = c(seq(1,12,1)), labels = month) +
 # scale_y_continuous(limits = c(1500, 3000), breaks = c(seq(1500,3000,250))) + 
#  geom_smooth(aes(x = week, y = probs/100), span = .5, se = FALSE, color = "gray70", size = 1) +
  theme_bw() +
  xlab("Year") +
  ylab("Maximum Sea Surface Temperatures in Caribbean [degrees C]") +
  facet_wrap(~ Year, ncol = 3)

yearSSTgraph <- yearSSTgraph + theme(axis.title.y = element_text(colour = "black"), axis.text.y = element_text(color = "black"),  axis.text.x = element_text(angle = 60, hjust = 1), axis.text = element_text(size = 10), axis.title=element_text(size=14), legend.title = element_blank(), legend.position = "bottom", legend.key.size = unit(1, "cm"),legend.text=element_text(size=12), panel.spacing = unit(1, "lines"))

yearSSTgraph
```

```{r}
 slopevals = numeric()
slopeargs = numeric()
for (i in 1994:2020) {
      dat <- ClimatebyYear %>% filter(Year == i)
      model = lm(ElNino_max ~ Mo, data = dat)
      slope = model$coefficients[2]
      slopearg = ifelse(model$coefficients[2] >= 0,  "POS", "NEG")
      slopeargs = append(slopeargs, slopearg)
      slopevals = append(slopevals, slope)
}

slope_group = character()
slope_vals = numeric()

for (i in 1:dim(ClimatebyYear)[1]) {
  yr = ClimatebyYear$Year[i]
  yearfordat = (yr - 1994)
  slope_group = append(slope_group, as.data.frame(slopeargs)[1 + yearfordat,])
  slope_vals = append(slope_vals, as.data.frame(slopevals)[1 + yearfordat,])
}

ClimatebyYear$group_ElNino <- slope_group
ClimatebyYear$slope_ElNino <- slope_vals
```

```{r}
month <- month.abb

yearElNinograph <- ggplot(ClimatebyYear, (aes(x = Mo, y = ElNino_max))) +
 geom_smooth(aes(color=group_ElNino), method = lm, span = .5, se = FALSE, size = 1) +
  scale_colour_manual(labels = c("Negative Trend", "Positive Trend"), values = c('red', 'blue')) +
  geom_line(color = "black", lwd = 1) +
  scale_x_continuous(expand = c(0, 0), breaks = c(seq(1,12,1)), labels = month) +
 # scale_y_continuous(limits = c(1500, 3000), breaks = c(seq(1500,3000,250))) + 
#  geom_smooth(aes(x = week, y = probs/100), span = .5, se = FALSE, color = "gray70", size = 1) +
  theme_bw() +
  xlab("Year") +
  ylab("Maximum El Nino Temperatures [degrees C]") +
  facet_wrap(~ Year, ncol = 3)

yearElNinograph <- yearElNinograph + theme(axis.title.y = element_text(colour = "black"), axis.text.y = element_text(color = "black"),  axis.text.x = element_text(angle = 60, hjust = 1), axis.text = element_text(size = 10), axis.title=element_text(size=14), legend.title = element_blank(), legend.position = "bottom", legend.key.size = unit(1, "cm"),legend.text=element_text(size=12), panel.spacing = unit(1, "lines"))

yearElNinograph
```

```{r}
 slopevals = numeric()
slopeargs = numeric()
for (i in 1994:2020) {
      dat <- ClimatebyYear %>% filter(Year == i)
      model = lm(monthlySIE_max ~ Mo, data = dat)
      slope = model$coefficients[2]
      slopearg = ifelse(model$coefficients[2] >= 0, "POS", "NEG")
      slopeargs = append(slopeargs, slopearg)
      slopevals = append(slopevals, slope)
}

slope_group = character()
slope_vals = numeric()

for (i in 1:dim(ClimatebyYear)[1]) {
  yr = ClimatebyYear$Year[i]
  yearfordat = (yr - 1994)
  slope_group = append(slope_group, as.data.frame(slopeargs)[1 + yearfordat,])
  slope_vals = append(slope_vals, as.data.frame(slopevals)[1 + yearfordat,])
}

ClimatebyYear$group_monthlySIE <- slope_group
ClimatebyYear$slope_monthlySIE <- slope_vals
```

```{r}
month <- month.abb

yearmonthlySIEgraph <- ggplot(ClimatebyYear, (aes(x = Mo, y = monthlySIE_max))) +
 geom_smooth(aes(color=group_monthlySIE), method = lm, span = .5, se = FALSE, size = 1) +
  scale_colour_manual(labels = c("Negative Trend", "Positive Trend"), values = c('red', 'blue')) +
  geom_line(color = "black", lwd = 1) +
  scale_x_continuous(expand = c(0, 0), breaks = c(seq(1,12,1)), labels = month) +
 # scale_y_continuous(limits = c(1500, 3000), breaks = c(seq(1500,3000,250))) + 
#  geom_smooth(aes(x = week, y = probs/100), span = .5, se = FALSE, color = "gray70", size = 1) +
  theme_bw() +
  xlab("Year") +
  ylab("Maximum Arctic Sea Ice Extent [millions of square kilometers]") +
  facet_wrap(~ Year, ncol = 3)

yearmonthlySIEgraph <- yearmonthlySIEgraph + theme(axis.title.y = element_text(colour = "black"), axis.text.y = element_text(color = "black"),  axis.text.x = element_text(angle = 60, hjust = 1), axis.text = element_text(size = 10), axis.title=element_text(size=14), legend.title = element_blank(), legend.position = "bottom", legend.key.size = unit(1, "cm"),legend.text=element_text(size=12), panel.spacing = unit(1, "lines"))

yearmonthlySIEgraph
```

*Yearly trends in climate variables*
```{r}
library(dplyr)
x <-  BigDays.sfdfT %>%
  group_by(Year) %>%
  summarize(maxmonthlySIE = max(monthlySIE))

monthlySIEgraph <- ggplot(x, (aes(x = Year, y = maxmonthlySIE))) +
 geom_smooth(method = lm, span = .5, se = FALSE, color = "red", size = 1) +
  geom_line(color = "black", lwd = 1) +
  scale_x_continuous(expand = c(0, 0), breaks = c(seq(1994,2020,1))) +
  scale_y_continuous(limits = c(14, 16), breaks = c(seq(14,16,0.25))) + 
#  geom_smooth(aes(x = week, y = probs/100), span = .5, se = FALSE, color = "gray70", size = 1) +
  theme_bw() +
  xlab("Year") +
  ylab("Maximum Sea Ice Extent [millions of square km]")

monthlySIEgraph <- monthlySIEgraph + theme(axis.title.y = element_text(colour = "black"), axis.text.y = element_text(color = "black"),  axis.text.x = element_text(angle = 45, hjust = 1), axis.text = element_text(size = 12), axis.title=element_text(size=14))
monthlySIEgraph
```

```{r}
x <-  BigDays.sfdfT %>%
  group_by(Year) %>%
  summarize(maxNino = max(monthlyElNino34))

Ninograph <- ggplot(x, (aes(x = Year, y = maxNino))) +
 geom_smooth(method = lm, span = .5, se = FALSE, color = "blue", size = 1) +
  geom_line(color = "black", lwd = 1) +
  scale_x_continuous(expand = c(0, 0), breaks = c(seq(1994,2020,1))) +
  scale_y_continuous(limits = c(26.5, 30), breaks = c(seq(26.5,30,0.5))) + 
#  geom_smooth(aes(x = week, y = probs/100), span = .5, se = FALSE, color = "gray70", size = 1) +
  theme_bw() +
  xlab("Year") +
  ylab(" Maximum El Nino 3.4 Index")

Ninograph <- Ninograph + theme(axis.title.y = element_text(colour = "black"), axis.text.y = element_text(color = "black"),  axis.text.x = element_text(angle = 45, hjust = 1), axis.text = element_text(size = 12), axis.title=element_text(size=14))
Ninograph
```

```{r}
x <-  BigDays.sfdfT %>%
  group_by(Year) %>%
  summarize(maxairtemp = max(avg_airtemp))

airtempgraph <- ggplot(x, (aes(x = Year, y = maxairtemp))) +
 geom_smooth(method = lm, span = .5, se = FALSE, color = "blue", size = 1) +
  geom_line(color = "black", lwd = 1) +
  scale_x_continuous(expand = c(0, 0), breaks = c(seq(1994,2020,1))) +
  scale_y_continuous(limits = c(0,1.5), breaks = c(seq(0,1.5,0.25))) + 
#  geom_smooth(aes(x = week, y = probs/100), span = .5, se = FALSE, color = "gray70", size = 1) +
  theme_bw() +
  xlab("Year") +
  ylab("Maximum Air Temperature Anomalies")

airtempgraph <- airtempgraph + theme(axis.title.y = element_text(colour = "black"), axis.text.y = element_text(color = "black"),  axis.text.x = element_text(angle = 45, hjust = 1), axis.text = element_text(size = 12), axis.title=element_text(size=14))
airtempgraph
```

```{r}
x <-  BigDays.sfdfT %>%
  group_by(Year) %>%
  summarize(maxSST = max(SSTgradient))

SSTgraph <- ggplot(x, (aes(x = Year, y = maxSST))) +
 geom_smooth(method = lm, span = .5, se = FALSE, color = "blue", size = 1) +
  geom_line(color = "black", lwd = 1) +
  scale_x_continuous(expand = c(0, 0), breaks = c(seq(1994,2020,1))) +
  scale_y_continuous(limits = c(34,50.2), breaks = c(seq(34,50,2))) + 
#  geom_smooth(aes(x = week, y = probs/100), span = .5, se = FALSE, color = "gray70", size = 1) +
  theme_bw() +
  xlab("Year") +
  ylab("Maximum SST gradient [degrees C]")

SSTgraph <- SSTgraph + theme(axis.title.y = element_text(colour = "black"), axis.text.y = element_text(color = "black"),  axis.text.x = element_text(angle = 45, hjust = 1), axis.text = element_text(size = 12), axis.title=element_text(size=14))
SSTgraph
```



```{r}
ggarrange(monthlySIEgraph, Ninograph, airtempgraph, SSTgraph, ncol = 2, nrow = 2)
```

## Descriptive Statistics

```{r}
x <- BigDays.sfdfT %>%
  filter(nT==10)
dim(x)[1]
#88 clusters have 10 tornadoes
x <- BigDays.sfdfT %>%
  filter(nT>=30)
dim(x)[1]
#166 clusters have at least 30 tornadoes
x <- BigDays.sfdfT %>%
  filter(nT>=50)
dim(x)[1]
#52 clusters have at least 50 tornadoes. 
```

```{r}
x <- BigDays.sfdfT %>%
  filter(GroupDayCas==0)
dim(x)[1]
#293 clusters have 0 casualties
x <- BigDays.sfdfT %>%
  filter(GroupDayCas>=30)
dim(x)[1]
#132 clusters have at least 30 casualties
x <- BigDays.sfdfT %>%
  filter(GroupDayCas>=50)
dim(x)[1]
#90 clusters have at least 50 casualties
```

#The Models: 
*What is the correlation between the climate variables? *
```{r}
ClimVars <- data.frame(AirTemperature = BigDays.sfdfT$avg_airtemp, ATL_SST = BigDays.sfdfT$monthlyATLSST, monthlyGAKSST = BigDays.sfdfT$monthlyGAKSST, ArcticmonthlySIE = BigDays.sfdfT$monthlySIE, ElNino = BigDays.sfdfT$monthlyElNino34)

xtable(cor(ClimVars))
```

*Create a figure of the correlation matrix. *
```{r}
library(corrplot)
names(ClimVars)[names(ClimVars) == 'AirTemperature'] <- 'AirTemp'

corrplot.mixed(cor(ClimVars), tl.col="black", lower.col="black", number.cex = 1, tl.cex = 1)

```



```{r}
(plot_CAPE <- ggplot(BigDays.sfdfT, aes(maxCAPE)) + 
  geom_histogram(boundary = 50, 
                 bins = 43, 
                 alpha = 0.9, 
                 color = "white", 
                 fill = "Grey40",
                 size = 0.1) +
  scale_x_continuous(expand = c(0,0), limits = c(0,7000)) +
  #scale_x_log10() + 
  xlab("Maximum CAPE [J/kg]") +
  scale_y_continuous(expand = c(0,0), breaks = c(0, 10,20,30,40,50), labels = c("0", "10", "20", "30", "40", "50"), limits = c(0, 50)) +
  ylab("Number of Clusters") + 
  ggtitle("A") +
  theme_bw() +
  theme(text = element_text(size=12), plot.background = element_rect(color = "grey40"), panel.border = element_rect(color = "white"), plot.margin=unit(c(0.5, 0.5, 0.5, 0.5),"cm")))
```

```{r}
(plot_DLBS <- ggplot(BigDays.sfdfT, aes(maxBS_deep)) + 
  geom_histogram(boundary = 50, 
                 bins = 43, 
                 alpha = 0.9, 
                 color = "white", 
                 fill = "Grey40",
                 size = 0.1) +
  scale_x_continuous(expand = c(0,0), limits = c(0,50)) +
  #scale_x_log10() + 
  xlab("Maximum Deep Layer Bulk Shear [m/s]") +
  scale_y_continuous(expand = c(0,0), breaks = c(0, 10,20,30,40,50, 60), labels = c("0", "10", "20", "30", "40", "50", "60"), limits = c(0, 60)) +
  ylab("Number of Clusters") + 
  ggtitle("C") +
  theme_bw() +
  theme(text = element_text(size=12), plot.background = element_rect(color = "grey40"), panel.border = element_rect(color = "white"), plot.margin=unit(c(0.5, 0.5, 0.5, 0.5),"cm")))
```

```{r}
(plot_SLBS <- ggplot(BigDays.sfdfT, aes(maxBS_shallow)) + 
  geom_histogram(boundary = 50, 
                 bins = 43, 
                 alpha = 0.9, 
                 color = "white", 
                 fill = "Grey40",
                 size = 0.1) +
  scale_x_continuous(expand = c(0,0), limits = c(0,40)) +
  #scale_x_log10() + 
  xlab("Maximum Shallow Layer Bulk Shear [m/s]") +
  scale_y_continuous(expand = c(0,0), breaks = c(0, 10,20,30,40,50), labels = c("0", "10", "20", "30", "40", "50"), limits = c(0, 50)) +
  ylab("Number of Clusters") + 
  ggtitle("D") +
  theme_bw() +
  theme(text = element_text(size=12), plot.background = element_rect(color = "grey40"), panel.border = element_rect(color = "white"), plot.margin=unit(c(0.5, 0.5, 0.5, 0.5),"cm")))
```

```{r}
(plot_CIN <- ggplot(BigDays.sfdfT, aes(minCIN)) + 
  geom_histogram(boundary = 50, 
                 bins = 43, 
                 alpha = 0.9, 
                 color = "white", 
                 fill = "Grey40",
                 size = 0.1) +
  scale_x_continuous(expand = c(0,0), limits = c(-700,50)) +
  #scale_x_log10() + 
  xlab("Minimum CIN [J/kg]") +
  scale_y_continuous(expand = c(0,0), breaks = c(0, 20,40,60,80,100), labels = c("0", "20", "40", "60", "80", "100"), limits = c(0, 100)) +
  ylab("Number of Clusters") + 
  ggtitle("B") +
  theme_bw() +
  theme(text = element_text(size=12), plot.background = element_rect(color = "grey40"), panel.border = element_rect(color = "white"), plot.margin=unit(c(0.5, 0.5, 0.5, 0.5),"cm")))
```


```{r}
ggarrange(plot_CAPE, plot_CIN, plot_DLBS, plot_SLBS, ncol = 2, nrow = 2)
```

```{r}
ggplot(BigDays.sfdfT, aes(x = Year, y = monthlySIE, color = Mo)) +
  geom_point()
ggplot(BigDays.sfdfT, aes(x = Year, y = avg_airtemp, color = Mo)) +
  geom_point()
ggplot(BigDays.sfdfT, aes(x = Year, y = monthlyATLSST, color = Mo)) +
  geom_point()
ggplot(BigDays.sfdfT, aes(x = Year, y = monthlyGAKSST, color = Mo)) +
  geom_point()
ggplot(BigDays.sfdfT, aes(x = Year, y = monthlyElNino34, color = Mo)) +
  geom_point()


BigDays.sfdfT %>%
  group_by(Year) %>%
  summarise(avgmonthlySIE = mean(monthlySIE)) %>%
ggplot(aes(x = Year, y = avgmonthlySIE)) +
  geom_point() +
  geom_smooth(method = lm)
```

```{r}
CAPE_m1 <- lmer(maxCAPE ~ Year + avg_airtemp + I(monthlyATLSST - monthlyGAKSST) + monthlySIE + monthlyElNino34 + (1|Mo), data = BigDays.sfdfT)
summary(CAPE_m1)
```
#Remove El Nino and AirTemp

```{r}
CAPE_m2 <- lmer(maxCAPE ~ Year + I(monthlyATLSST - monthlyGAKSST) + monthlySIE + (1|Mo), data = BigDays.sfdfT)
summary(CAPE_m2)
ranef(CAPE_m2)
```

```{r}
CAPE_m3 <- lmer(maxCAPE ~ I(monthlyATLSST - monthlyGAKSST) + (1|Mo) + (1|Year), data = BigDays.sfdfT)
summary(CAPE_m3)

hist(BigDays.sfdfT$monthlySIE)

18571/33143 #Check terms by EF1+,2,3,...
```