---
title: "Update_BigDays"
author: "Zoe Schroder"
date: "12/9/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

The data for the BigDays needs to be updated. Currently it only contains information from 1994 to 2018. This needs to be extended to 2020. As such, additional NARR data and climate data will need to be downloaded and added to the `RData` file. This work will closely follow that outlined in the **DefineBigDays** repository on my github page. 

The 1950 to 2018 tornado data is downloaded as a csv file from https://www.spc.noaa.gov/wcm/#data

The 2019 through September 2020 data came from the SPC National Center for Environmental Information Storm Events Database (https://www.ncdc.noaa.gov/stormevents/choosedates.jsp?statefips=-999%2CALL). 

*Load the necessary packages*
```{r}
suppressMessages(library(sp))
suppressMessages(library(sf))
suppressMessages(library(lubridate))
suppressMessages(library(dplyr))
#install.packages("chron")
suppressMessages(library(chron))
suppressMessages(library(raster))
library(USAboundaries)
library(tmap)
library(ggplot2)
```

Projections you may need: 
```{r}
merc <- "+proj=merc +lon_0=0 +k=1 +x_0=0 +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"
US_LCC <- "+proj=lcc +lat_1=33 +lat_2=45 +lat_0=39 +lon_0=-96 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m no_defs"
```

Coordinate Reference Systems you may need:
```{r, eval = FALSE}
WGS84 <- 4326
```


Use the start lon/lat and create a `sp` object then convert to `sf`. Set the coordinate reference system (crs) to ESPG to `WGS84`.
```{r, eval = FALSE}
Torn50_19.spdf <- read.csv(file = "1950-2019_actual_tornadoes.csv")
sp::coordinates(Torn50_19.spdf) <- ~ slon + slat
Torn50_19.sfdf <- st_as_sf(Torn50_19.spdf)
st_crs(Torn50_19.sfdf) <- WGS84
```

```{r, eval = FALSE}
#Torn50_18.spdf <- read.csv(file = "1950-2018_actual_tornadoes.csv")
#sp::coordinates(Torn50_18.spdf) <- ~ slon + slat
#Torn50_18.sfdf <- st_as_sf(Torn50_18.spdf)
#st_crs(Torn50_18.sfdf) <- WGS84
```

```{r, eval = FALSE}
#timecheck <- Torn50_19.sfdf %>%
#  filter(date == Torn50_18.sfdf$date && (Torn50_19.sfdf$time %in% Torn50_18.sfdf$time))
```

In the 2020 data, we can also use the start lat and lon to create an `sp` object. We then can convert it to `sf` data frame and set the coordinate reference system (crs) to EPSG 4326 [WGS 84]. However, we need to adjust the column headers of the 2020 tornado files to match those of the 1950 to 2019 data frame. 

Read the 2020 tornado data into R.
```{r, eval = FALSE}
Torn2020.spdf <-read.csv(file = "2020_TornadoData.csv")
```

The column headers do not match the 1950 to 2019 data frame. Therefore, you MUST rename the column headers before you can combine the data frames. Change the column names using the `colnames` function. 
```{r, eval = FALSE}
cols <- c("eventid", "county", "city", "date", "originaltime", "eventtype", "mag_NA", "mag", "fat", "inj", "loss", "closs", "st", "timezone", "magtype", "epiid", "cztype", "stf", "WFO", "injind", "fatind", "source", "floodcause", "len", "wid", "brange", "bazimuth", "erange", "eazimuth", "elocation", "slat", "slon", "elat", "elon", "eventnar", "episodenar", "om")
colnames(Torn2020.spdf) <- cols
```

Convert to a spatial points data frame using the start latitude and longitude for each tornado. Next, convert to a simple features data frame with a set coordinate reference system to the WGS 84 (`EPSG: 4236`)
```{r, eval = FALSE}
sp::coordinates(Torn2020.spdf) <- ~ slon + slat
Torn2020.sfdf <- st_as_sf(Torn2020.spdf)
st_crs(Torn2020.sfdf) <- WGS84
```

Adjust the date and time columns. Add columns for day (dy), month (mo), year (yr), time(CST) 
```{r, eval = FALSE}
mydates <- as.Date(c(Torn2020.sfdf$date), format = "%m/%d/%Y")

Torn2020.sfdf <-  Torn2020.sfdf %>%
  mutate(yr = rep(2020, 1131),
         dy = day(as.Date(c(date), format = "%m/%d/%Y")),
         mo = month(as.Date(c(date), format = "%m/%d/%Y")),
         mag = as.integer(substr(mag, 3,3)))
```

Convert times to the CST time zone. In the originaltime column, it has values of CST-6, EST-5, MST-7, and PST-8. Therefore, we create a `for` loop to convert these times to CST which is consistent with `Torn50_19.sfdf`.
```{r, eval = FALSE}
time <- c()
date <- c()

Torn2020.sfdf$date <- as.Date(Torn2020.sfdf$date, format = "%m/%d/%Y")

for (i in 1:1131){
  if(Torn2020.sfdf$timezone[i] == "CST-6") { 
    time <- append(time,Torn2020.sfdf$originaltime[i])
    date <- append(date, as.Date(Torn2020.sfdf$date[i]))
    } else if(Torn2020.sfdf$timezone[i] == "EST-5") {
    o_time <- ifelse((Torn2020.sfdf$originaltime[i] <= 100), (Torn2020.sfdf$originaltime[i] +1200-100), Torn2020.sfdf$originaltime[i] - 100)
    newdate <- if(Torn2020.sfdf$originaltime[i] <= 100) {as.Date(Torn2020.sfdf$date[i], format = "%m/%d/%Y", origin = "0020/01/01")-1 } else { as.Date(Torn2020.sfdf$date[i], format = "%m/%d/%Y")
    }
    date <- append(date, newdate)
    time <- append(time, (o_time))  
    }   else if(Torn2020.sfdf$timezone[i] == "MST-7") {
    o_time <- ifelse((Torn2020.sfdf$originaltime[i] >= 2300), (Torn2020.sfdf$originaltime[i]  - 2300), Torn2020.sfdf$originaltime[i] +100)
    newdate <- if(Torn2020.sfdf$originaltime[i] >= 2300) {as.Date(Torn2020.sfdf$date[i], format = "%m/%d/%Y", origin = "0020/01/01")+1 } else { as.Date(Torn2020.sfdf$date[i], format = "%m/%d/%Y")
    }
    time <- append(time, Torn2020.sfdf$originaltime[i])
    date <- append(date, as.Date(Torn2020.sfdf$date[i])) 
    }    else if(Torn2020.sfdf$timezone[i] == "PST-8") {
          o_time <- ifelse((Torn2020.sfdf$originaltime[i] >= 2200), (Torn2020.sfdf$originaltime[i]  - 2200), Torn2020.sfdf$originaltime[i] +200)
    newdate <- if(Torn2020.sfdf$originaltime[i] >= 2200) {as.Date(Torn2020.sfdf$date[i], format = "%m/%d/%Y", origin = "0020/01/01")+1 } else { as.Date(Torn2020.sfdf$date[i], format = "%m/%d/%Y")
      }
    time <- append(time,(Torn2020.sfdf$originaltime[i]))
    date <- append(date, as.Date(Torn2020.sfdf$date[i]))
    } else ("NA")
}

time <- chron(times=time)
time <- format(strptime(substr(as.POSIXct(sprintf("%04.0f", time), format="%H%M"), 12, 16), '%H:%M'), '%I:%M:%S %p')
time <- format(strptime(time, "%I:%M:%S"), "%H:%M:%S")

x<-as.POSIXct(time,format="%H:%M:%S")
hour <- hour(x)
tz <- rep(3, 1131)

Torn2020.sfdf <- cbind(Torn2020.sfdf, time, hour, tz)
```

Determine the column names in the `Torn50_19.sfdf`. 
```{r, eval = FALSE}
colnames(Torn50_19.sfdf)
```

Determine the column names in the `Torn2020.sfdf`. 
```{r, eval = FALSE}
colnames(Torn2020.sfdf)
```

Filter the column names in the `Torn50_19.sfdf` and `Torn2020.sfdf`. You want both data frames to have the same columns.  
```{r, eval = FALSE}
filter_cols <- c("om","yr", "mo", "dy", "date", "time", "tz", "st", "stf", "mag","inj", "fat", "loss", "closs", "elat", "elon", "len", "wid", "geometry")

Torn50_19.sfdf <-  Torn50_19.sfdf %>%
  dplyr::select(filter_cols)

Torn2020.sfdf <-  Torn2020.sfdf %>%
  dplyr::select(filter_cols)
#stn sn, ns, sg, f1m f2, f3, fc were removed from the 1950:2019 data file. 
```

Combine the 2020 and 1950 to 2019 data frames.
```{r, eval = FALSE}
Torn.sfdf <- as.data.frame(rbind(Torn50_19.sfdf, Torn2020.sfdf))
```

Remove tornadoes in Hawaii, Alaska, and Puerto Rico and those occurring before 1994. That year marks the beginning of comprehensive WSR-88D radar. For missing EF ratings use the modification rules (if/else) defined here: https://www.spc.noaa.gov/wcm/OneTor_F-scale-modifications.pdf 
```{r, eval = FALSE}
All_Tornadoes <- Torn.sfdf %>%
  filter(yr >= 1994,
         !st %in% c("AK", "PR", "HI")) %>%
  mutate(mag = ifelse(mag == -9 & len <= 5, 0, mag),
         mag = ifelse(mag == -9 & len > 5, 1, mag))
```

Add a data/time column also add columns for path length, width, and area in metric units. Leave the time zone as native CDT. Create a convective day (6AM to 6AM) column taking hours 00:00:00 -> 05:59:59 and assigning it to the previous date (this associates the previous day's date to tornadoes occurring up to 6 hours after local midnight).
```{r, eval = FALSE}
All_Tornadoes <- All_Tornadoes %>%
  mutate(#dy = format(as.Date(date,format="%y/%m/%d"), "%d"), #This is now an included column in data
         DateTime = as.POSIXct(paste(yr, mo, dy, time), format = "%Y%m%d%H:%M:%S"),
         Hour = hour(DateTime),
         Year = year(DateTime),
         cDateTime = DateTime - as.difftime(6, unit = "hours"),
         cDate = as.Date(as_datetime(ifelse(Hour < 6, (DateTime - 86400), cDateTime), tz = Sys.timezone())),
         Length = len * 1609.34,
         Length = ifelse(Length == 0, min(Length[Length > 0]), Length), #takes care of zero length
         Width = wid * .9144,
         Width = ifelse(Width == 0, min(Width[Width > 0]), Width), #takes care of zero width
         Width = ifelse(Year >= 1995, Width * pi/4, Width), #takes care of change: avg to max
         cas = inj + fat,
         AreaPath = Length * Width,
         Ma = factor(month.abb[mo], levels = month.abb[1:12])) %>%
  sf::st_sf()
dim(All_Tornadoes)[1]
```

The geometry type is `POINT`. Each tornado is represented as a single point location geometry (start location). Add power dissipation per tornado.

Add power dissipation per tornado. Use the empirical model for tornado winds by EF rating taken from Table 3-1 of NRC 2007. Percent area by EF rating for each EF category. Threshold wind speeds (m/s) are a lower bound 3-sec gusts on the operational EF Scale (Table 2-1 of NRC2007). This is based on work by Fricker et al. (2017). The model is
$$
E = A_p \rho \sum_{j=0}^{J} w_j v_j^{3},
$$
where $A_p$ is the area of the path, $\rho$ is area density [1 kg/m^3]  $v_j$ is the midpoint wind speed for each rating, and $w_j$ is the corresponding fraction of path area by EF rating. With no upper bound on the EF5 wind speeds, the midpoint wind speed is set at 97 m~s$^{-1}$ (7.5 m~s$^{-1}$ above the threshold wind speed consistent with the EF4 midpoint speed relative to its threshold)
```{r, eval = FALSE}
perc <- c(1, 0, 0, 0, 0, 0, 
          .772, .228, 0, 0, 0, 0,
          .616, .268, .115, 0, 0, 0,
          .529, .271, .133, .067, 0, 0,
          .543, .238, .131, .056, .032, 0,
          .538, .223, .119, .07, .033, .017)
percM <- matrix(perc, ncol = 6, byrow = TRUE)
threshW <- c(29.06, 38.45, 49.62, 60.8, 74.21, 89.41)
midptW <- c(diff(threshW)/2 + threshW[-length(threshW)], threshW[length(threshW)] + 7.5)
ef <- All_Tornadoes$mag + 1
EW3 <- numeric()
for(i in 1:length(ef)) EW3[i] = midptW^3 %*% percM[ef[i], ]
All_Tornadoes <- All_Tornadoes %>%
  mutate(ED = EW3 * AreaPath)
``` 

Determine the distance between tornadoes in space and time. Use a projection, not lat/lon. See https://epsg.io/102004. Extract the coordinates of the start locations as a N by 2 matrix, where N is the number of tornadoes. Also extract the date-time as a vector of class `POSIXct`.
```{r, eval = FALSE}
All_Tornadoes <- st_transform(All_Tornadoes, crs = st_crs(US_LCC))
space <- st_coordinates(All_Tornadoes)
time <- All_Tornadoes$DateTime
```

Next compute pairwise Euclidean distances in space and, separately, in time using the `dist()` function. Divide the spatial distance by 15 so that the values are commensurate with the time 'distance' based on the assumption of 15 meters per second (~34 mph) for an average speed of tornado-generating storms. Compare: Distance from New York to Denver is 2.622 x 10^6 meters. There are 3.154 x 10^7 seconds in a year. This will capture the historic multiday tornado outbreaks. For analysis we want to consider each day in the multiday group separately. As the value of the divisor increases cluster areas get larger. Remove `ds` and `dt` to free memory. Distances are saved as an object of class `dist` containing a vector of length N * (N-1)/2, which is the number of unique point pairs.
```{r, eval = FALSE}
ds <- (dist(space) / 15)
dt <- dist(time)
dst <- ds + dt
rm(ds, dt)
```

Distances are saved as an object of class `dist` containing a vector of length N * (N-1)/2, which is the number of unique point pairs.

Next group the tornadoes based on the space-time distances. This is done with the `hclust()` (hierarchical cluster) function. Initially, each tornado is assigned to its own group and then the algorithm joins the two closest tornadoes determined by values in `dst`. The algorithm continues by joining tornadoes (and tornado groups) until there is a single large group.

The single linkage method (`method = "single"`) is related to the minimal spanning tree (MST) and adopts a 'friends of friends' grouping strategy. An edge-weighted graph is a graph where each edge has a weight (or cost). Here weights are space-time distances between tornadoes. A MST of an edge-weighted graph is a spanning tree whose weight (the sum of the weights of its edges) is no larger than the weight of any other spanning tree. A spanning tree of a graph on N vertices (tornado centroids) is a subset of N-1 edges that form a tree (Skiena 1990, p. 227).
 
The `cutree()` function is used to extract a group number for each tornado. Tornadoes in each group are close in space & time. Here the tree is cut at a height of 50000 space-time units. Making `h` smaller results in smaller groups (fewer tornadoes per group).
```{r, eval = FALSE}
stime <- proc.time()
tree <- hclust(dst, method = "single")
groupNumber <- as.integer(cutree(tree, h = 50000))
proc.time() - stime
```

Add the group number to each tornado. 
```{r, eval = FALSE}
All_Tornadoes$groupNumber <- groupNumber
```

Compute group-level statistics. 
```{r, eval = FALSE}
Groups.sfdfT <- All_Tornadoes %>%
  group_by(groupNumber) %>%
  summarize(Year = first(Year),
            Month = first(mo),
            FirstDate = first(date),
            LastDate = last(date),
            Name = paste(FirstDate, "to", LastDate),
            FirstcDate = first(cDate),
            LastcDate = last(cDate),
            ncD = n_distinct(cDate),
            nT = n(),
            n0 = sum(mag == 0),
            n1 = sum(mag == 1),
            n2 = sum(mag == 2),
            n3 = sum(mag == 3),
            n4 = sum(mag == 4),
            n5 = sum(mag == 5),
            ATP = sum(ED),
            ATP_TW = paste(round(ATP/10^12), "TW"),
            maxEF = max(mag),
            nD = n_distinct(date),
            StartTime = first(DateTime),
            EndTime = last(DateTime),
            Duration = difftime(EndTime, StartTime, units = "secs"), 
            cas = sum(inj + fat)) 
```

######################
## Extract Big Days ##
######################

Filter individual tornadoes to remove those that are not part of a large group. Group by group number and convective dates. Remove days having fewer than 10 tornadoes.
```{r, eval = FALSE}
BigDays.sfdfT <- All_Tornadoes %>%
  filter(groupNumber %in% Groups.sfdfT$groupNumber) %>%
  group_by(groupNumber, cDate) %>%
  summarize(nT = n(),
            n0 = sum(mag == 0),
            n1 = sum(mag == 1),
            n2 = sum(mag == 2),
            n3 = sum(mag == 3),
            n4 = sum(mag == 4),
            n5 = sum(mag == 5),
            ATP = sum(ED),
            maxATP = max(ED),
            avgATP = mean(ED),
            GroupDayCas = sum(cas),
            GroupDayFat = sum(fat),
            StartTime_CST = first(DateTime),
            EndTime_CST= last(DateTime),
            StartTime_UTC = StartTime_CST + 21600,
            EndTime_UTC = EndTime_CST + 21600,
            Duration = difftime(EndTime_CST, StartTime_CST, units = "secs")) %>%
  filter(nT >= 10) %>%
  mutate(Year = year(cDate),
         Mo = month(cDate),
         Month = format(cDate, "%m"), # this is needed to preserve the leading zeros
         Day = format(cDate, "%d"), 
         ATP_TW = ATP/10^12)                                                                              
dim(BigDays.sfdfT)
```

Create a unique ID for each big day and each tornado. Extract the tornadoes associated with each big day using the unique ID.
```{r, eval = FALSE}
BigDayTornadoes <- All_Tornadoes %>%
   mutate(ID = paste0(gsub("-", "", cDate), groupNumber))
BigDays.sfdfT <- BigDays.sfdfT %>%
   mutate(ID = paste0(gsub("-", "", cDate), groupNumber))

BigDayTornadoes <- BigDayTornadoes %>%
  filter(ID %in% BigDays.sfdfT$ID)

sum(BigDays.sfdfT$nT)
```

Round the UTC time to nearest 6 hours. This is done with the `align.time()` function from the **xts** package. Adjust it by 3 hours to get the closest time. This falls within the outbreak so you need to subtract by 3 hours (10800 seconds). This will produce the closest 3 hour NARR time that occurs before and not within the big day. 
```{r, eval = FALSE}
library(xts)
BigDays.sfdfT$StartTime_UTC <- force_tz(BigDays.sfdfT$StartTime_UTC, tzone = "UTC")
BigDays.sfdfT$NARRtime <- (align.time(BigDays.sfdfT$StartTime_UTC, n = (60 * 60 * 3)) - 3600 * 3)
```

Split the NARR date and time into their individual variables. Then bind the columns for BigDays.sfdfT. NOTE: cannot do a mutate because 00Z produces NAs. DON'T USE!
```{r, eval = FALSE}
NARRday = format(as.POSIXct(strptime(BigDays.sfdfT$NARRtime,"%Y-%m-%d %H:%M:%S",tz="")) ,format = "%Y/%m/%d")
NARRZtime = format(as.POSIXct(strptime(BigDays.sfdfT$NARRtime,"%Y-%m-%d %H:%M:%S",tz="")) ,format = "%H")

BigDays.sfdfT <- cbind(BigDays.sfdfT, NARRday, NARRZtime)
```

Create a downloadable string of information for the varying NARR times. 
```{r, eval = FALSE}
BigDays.sfdfT <- BigDays.sfdfT %>%
  mutate(YrMoDa = gsub("/", "", NARRday),
         slug = paste0("merged_AWIP32.",YrMoDa, NARRZtime),
         slug2 = paste0("merged_AWIP32.",YrMoDa))
```

Extract a vector of the big days. Save as a .csv for NARR download. 
```{r, eval = FALSE}
bigdays <- BigDays.sfdfT$NARRday
bigdaytimes <- BigDays.sfdfT$NARRZtime
x <- cbind(as.character(bigdays), as.character(bigdaytimes))
write.csv(x, "BigDays.csv")
```

```{r, eval = FALSE}
BigDays.sfdfT <- st_transform(BigDays.sfdfT, 
  crs = st_crs(US_LCC))
```

Obtain the group day hulls. Transform the CRS to match that of the environmental data raster grids.
```{r, eval = FALSE}
BigDays.sfdfT <- st_convex_hull(BigDays.sfdfT)
BigDays.sfdfT$HullArea <- st_area(BigDays.sfdfT)
BigDays.sfdfT <- st_transform(BigDays.sfdfT, 
  crs = "+proj=lcc +lat_1=50 +lat_2=50 +lat_0=50 +lon_0=-107 +x_0=0 +y_0=0 +a=6371200 +b=6371200 +units=m +no_defs")
```

Get the centroid (central point of the tornado activity) for each big day. 
```{r, eval = FALSE}
BigDayCentroids.df <- st_centroid(BigDays.sfdfT)
BigDayCentroids.df$groupArea <- st_area(st_convex_hull(BigDays.sfdfT))
BigDayCentroids.df$groupDensity <- BigDayCentroids.df$nT/BigDayCentroids.df$groupArea
```

Save the necessary data frames. 
```{r, eval = FALSE}
#save(All_Tornadoes, BigDayCentroids.df, BigDays.sfdfT, BigDayTornadoes, Groups.sfdfT, file = "BigDays.RData")
```

## Download NARR data: 

Data is downloaded from NCAR's North American Regional Reanalysis (https://rda.ucar.edu/datasets/ds608.0/#!access). It extends from 1-1-1979 to 11-1-2018. Use the NCAR NARR 3-hourly files.  

Spatial Extent: 
Longitude Range: Westernmost = 148.64E Easternmost = 2.568W
Latitude Range: Southernmost = 0.897N Northernmost = 85.333N

```{r, eval = FALSE}
BigDays.sfdfT <- st_transform(BigDays.sfdfT, 
  crs = "+proj=lcc +lat_1=50 +lat_2=50 +lat_0=50 +lon_0=-107 +x_0=0 +y_0=0 +a=6371200 +b=6371200 +units=m +no_defs")
```

The list of all variables can be found here: http://www.emc.ncep.noaa.gov/mmb/rreanl/merged_land_AWIP32.pdf 
```{r, eval = FALSE}
slug <- BigDays.sfdfT$slug
slug2 <- BigDays.sfdfT$slug2
```

Read the grib files as raster bricks and assign the CAPE and helicity variables to separate raster layers. Extract the average (and extreme) environmental values within each of the big days in large groups hulls.
```{r, eval = FALSE}
avgCAPE <- numeric()
avgsbCAPE <- numeric()
avgDEW <- numeric()
avgMR <- numeric()
avgHLCY <- numeric()
avgCIN <- numeric()
avgsbCIN <- numeric()
avgUSTM <- numeric()
avgVSTM <- numeric()
avgBS_deep <- numeric()
avgBS_shallow <- numeric()
avgSM <- numeric()
avgRATIO <- numeric()
avgLCL <- numeric()
maxCAPE <- numeric()
maxsbCAPE <- numeric()
maxDEW <- numeric()
maxMR <- numeric()
maxHLCY <- numeric()
minCIN <- numeric()
minsbCIN <- numeric()
maxUSTM <- numeric()
maxVSTM <- numeric()
maxBS_deep <- numeric()
maxBS_shallow <- numeric()
maxSM <- numeric()
maxLCL <- numeric()
minLCL <- numeric()
 
for(i in 1:length(slug)){
  print(i)
  rb <- brick(paste0("/Volumes/Zoe's Home/NCARNARR/All/", BigDays.sfdfT$slug2[i], "/",BigDays.sfdfT$slug[i])) #<-- this is for varying NARR times
  CAPE <- raster(rb, layer = 375)
  sbCAPE <- raster(rb, layer = 315)
  DEW <- raster(rb, layer = 290)
  MR <- raster(rb, layer = 289)
  HLCY <- raster(rb, layer = 323)
  CIN <- raster(rb, layer = 376)
  sbCIN <- raster(rb, layer = 316)
  USTM <- raster(rb, layer = 324)
  VSTM <- raster(rb, layer = 325)
  UGRD500 <- raster(rb, layer = 117) 
  VGRD500 <- raster(rb, layer = 118) 
  UGRD850 <- raster(rb, layer = 206) 
  VGRD850 <- raster(rb, layer = 207)
  UGRDsfc <- raster(rb, layer = 293) 
  VGRDsfc <- raster(rb, layer = 294)  
  LCL <- raster(rb, layer = 319)
  SM <- sqrt(USTM^2 + VSTM^2)
  RATIO <- CAPE/abs(CIN)
  BS_deep <- sqrt(((UGRD500 - UGRDsfc)**2) + ((VGRD500 - VGRDsfc)**2))
  BS_shallow <- sqrt(((UGRD850 - UGRDsfc)**2) + ((VGRD850 - VGRDsfc)**2))
  avgCAPE <- c(avgCAPE, as.numeric(raster::extract(CAPE, BigDays.sfdfT[i, ], fun = mean)))
  avgsbCAPE <- c(avgsbCAPE, as.numeric(raster::extract(sbCAPE, BigDays.sfdfT[i, ], fun = mean)))
  maxCAPE <- c(maxCAPE, as.numeric(raster::extract(CAPE, BigDays.sfdfT[i, ], fun = max)))
  maxsbCAPE <- c(maxsbCAPE, as.numeric(raster::extract(sbCAPE, BigDays.sfdfT[i, ], fun = max)))
  avgDEW <- c(avgDEW, as.numeric(raster::extract(DEW, BigDays.sfdfT[i, ], fun = mean)))
  maxDEW <- c(maxDEW, as.numeric(raster::extract(DEW, BigDays.sfdfT[i, ], fun = max)))
  avgMR <- c(avgMR, as.numeric(raster::extract(MR, BigDays.sfdfT[i, ], fun = mean)))
  maxMR <- c(maxMR, as.numeric(raster::extract(MR, BigDays.sfdfT[i, ], fun = max)))
  avgHLCY <- c(avgHLCY, as.numeric(raster::extract(HLCY, BigDays.sfdfT[i, ], fun = mean)))
  maxHLCY <- c(maxHLCY, as.numeric(raster::extract(HLCY, BigDays.sfdfT[i, ], fun = max)))
  avgCIN <- c(avgCIN, as.numeric(raster::extract(CIN, BigDays.sfdfT[i, ], fun = mean)))
  avgsbCIN <- c(avgsbCIN, as.numeric(raster::extract(sbCIN, BigDays.sfdfT[i, ], fun = mean)))
  minCIN <- c(minCIN, as.numeric(raster::extract(CIN, BigDays.sfdfT[i, ], fun = min)))
  minsbCIN <- c(minsbCIN, as.numeric(raster::extract(sbCIN, BigDays.sfdfT[i, ], fun = min)))
  avgUSTM <- c(avgUSTM, as.numeric(raster::extract(USTM, BigDays.sfdfT[i, ], fun = mean)))
  maxUSTM <- c(maxUSTM, as.numeric(raster::extract(USTM, BigDays.sfdfT[i, ], fun = max)))
  avgVSTM <- c(avgVSTM, as.numeric(raster::extract(VSTM, BigDays.sfdfT[i, ], fun = mean)))
  maxVSTM <- c(maxVSTM, as.numeric(raster::extract(VSTM, BigDays.sfdfT[i, ], fun = max)))
  avgSM <- c(avgSM, as.numeric(raster::extract(SM, BigDays.sfdfT[i, ], fun = mean)))
  maxSM <- c(maxSM, as.numeric(raster::extract(SM, BigDays.sfdfT[i, ], fun = max)))
  avgRATIO <- c(avgRATIO, as.numeric(raster::extract(RATIO, BigDays.sfdfT[i, ], fun = mean)))
  avgBS_deep <- c(avgBS_deep, as.numeric(raster::extract(BS_deep, BigDays.sfdfT[i, ], fun = mean)))
  maxBS_deep <- c(maxBS_deep, as.numeric(raster::extract(BS_deep, BigDays.sfdfT[i, ], fun = max)))  
  avgBS_shallow <- c(avgBS_shallow, as.numeric(raster::extract(BS_shallow, BigDays.sfdfT[i, ], fun = mean)))
  maxBS_shallow <- c(maxBS_shallow, as.numeric(raster::extract(BS_shallow, BigDays.sfdfT[i, ], fun = max)))
  avgLCL <- c(avgLCL, as.numeric(raster::extract(LCL, BigDays.sfdfT[i,], fun = mean)))
  maxLCL <- c(maxLCL, as.numeric(raster::extract(LCL, BigDays.sfdfT[i,], fun = max)))
  minLCL <- c(minLCL, as.numeric(raster::extract(LCL, BigDays.sfdfT[i,], fun = min)))
}
```

If you need to run your NARR data code in multiple sections, you can combine them using the chunk below. 
```{r, eval = FALSE}
avgCAPE811 <- avgCAPE
avgsbCAPE811 <- avgsbCAPE 
avgDEW811 <-  avgDEW 
avgMR811 <-  avgMR 
avgHLCY811 <- avgHLCY 
avgCIN811 <- avgCIN
avgsbCIN811 <-  avgsbCIN
avgUSTM811 <-  avgUSTM
avgVSTM811 <-  avgVSTM 
avgBS_deep811 <-  avgBS_deep
avgBS_shallow811 <- avgBS_shallow
avgSM811 <- avgSM
avgRATIO811 <-  avgRATIO
avgLCL811 <-  avgLCL
maxCAPE811 <-  maxCAPE
maxsbCAPE811 <-  maxsbCAPE
maxDEW811 <-  maxDEW 
maxMR811 <-  maxMR 
maxHLCY811 <-   maxHLCY
minCIN811 <-  minCIN
minsbCIN811 <-  minsbCIN 
maxUSTM811 <-  maxUSTM 
maxVSTM811 <-  maxVSTM 
maxBS_deep811 <-  maxBS_deep
maxBS_shallow811 <-  maxBS_shallow
maxSM811 <-  maxSM
maxLCL811 <-  maxLCL
minLCL811 <-  minLCL

avgCAPE830 <- avgCAPE
avgsbCAPE830 <- avgsbCAPE 
avgDEW830 <-  avgDEW 
avgMR830 <-  avgMR 
avgHLCY830 <- avgHLCY 
avgCIN830 <- avgCIN
avgsbCIN830 <-  avgsbCIN
avgUSTM830 <-  avgUSTM
avgVSTM830 <-  avgVSTM 
avgBS_deep830 <-  avgBS_deep
avgBS_shallow830 <- avgBS_shallow
avgSM830 <- avgSM
avgRATIO830 <-  avgRATIO
avgLCL830 <-  avgLCL
maxCAPE830 <-  maxCAPE
maxsbCAPE830 <-  maxsbCAPE
maxDEW830 <-  maxDEW 
maxMR830 <-  maxMR 
maxHLCY830 <-   maxHLCY
minCIN830 <-  minCIN
minsbCIN830 <-  minsbCIN 
maxUSTM830 <-  maxUSTM 
maxVSTM830 <-  maxVSTM 
maxBS_deep830 <-  maxBS_deep
maxBS_shallow830 <-  maxBS_shallow
maxSM830 <-  maxSM
maxLCL830 <-  maxLCL
minLCL830 <-  minLCL

avgCAPE <- append(avgCAPE811, avgCAPE830) 
avgsbCAPE<- append(avgsbCAPE811, avgsbCAPE830) 
avgDEW<- append(avgDEW811, avgDEW830) 
avgMR<- append(avgMR811, avgMR830) 
avgHLCY<- append(avgHLCY811, avgHLCY830) 
avgCIN<- append(avgCIN811, avgCIN830) 
avgsbCIN<- append(avgsbCIN811, avgsbCIN830) 
avgUSTM<- append(avgUSTM811, avgUSTM830) 
avgVSTM<- append(avgVSTM811, avgVSTM830) 
avgBS_deep<- append(avgBS_deep811, avgBS_deep830) 
avgBS_shallow<- append(avgBS_shallow811, avgBS_shallow830) 
avgSM<- append(avgSM811, avgSM830) 
avgRATIO<- append(avgRATIO811, avgRATIO830) 
avgLCL<- append(avgLCL811, avgLCL830) 
maxCAPE<- append(maxCAPE811, maxCAPE830) 
maxsbCAPE<- append(maxsbCAPE811, maxsbCAPE830) 
maxDEW<- append(maxDEW811, maxDEW830) 
maxMR<- append(maxMR811, maxMR830) 
maxHLCY<- append(maxHLCY811, maxHLCY830) 
minCIN<- append(minCIN811, minCIN830) 
minsbCIN<- append(minsbCIN811, minsbCIN830) 
maxUSTM<- append(maxUSTM811, maxUSTM830) 
maxVSTM<- append(maxVSTM811, maxCAPE830) 
maxBS_deep<- append(maxBS_deep811, maxBS_deep830) 
maxBS_shallow<- append(maxBS_shallow811, maxBS_shallow830) 
maxSM<- append(maxSM811, maxSM830) 
maxLCL<- append(maxLCL811, maxLCL830) 
minLCL<- append(minLCL811, minLCL830)
```

Add environmental data values to the group day means data frame.
```{r, eval = FALSE}
BigDays.sfdfT$avgCAPE <- avgCAPE
BigDays.sfdfT$avgsbCAPE <- avgsbCAPE
BigDays.sfdfT$maxCAPE <- maxCAPE
BigDays.sfdfT$maxsbCAPE <- maxsbCAPE
BigDays.sfdfT$avgDEW <- avgDEW
BigDays.sfdfT$maxDEW <- maxDEW
BigDays.sfdfT$avgMR <- avgMR
BigDays.sfdfT$maxMR <- maxMR
BigDays.sfdfT$avgHLCY <- avgHLCY
BigDays.sfdfT$maxHLCY <- maxHLCY
BigDays.sfdfT$avgCIN <- avgCIN
BigDays.sfdfT$avgsbCIN <- avgsbCIN
BigDays.sfdfT$minCIN <- minCIN
BigDays.sfdfT$minsbCIN <- minsbCIN
BigDays.sfdfT$avgUSTM <- avgUSTM
BigDays.sfdfT$maxUSTM <- maxUSTM
BigDays.sfdfT$avgVSTM <- avgVSTM
BigDays.sfdfT$maxVSTM <- maxVSTM
BigDays.sfdfT$avgBS_deep <- avgBS_deep
BigDays.sfdfT$maxBS_deep <- maxBS_deep
BigDays.sfdfT$avgBS_shallow <- avgBS_shallow
BigDays.sfdfT$maxBS_shallow <- maxBS_shallow
BigDays.sfdfT$avgRATIO <- avgRATIO
BigDays.sfdfT$avgSM <- avgSM
BigDays.sfdfT$maxSM <- maxSM
BigDays.sfdfT$minLCL <- minLCL
BigDays.sfdfT$maxLCL <- maxLCL
BigDays.sfdfT$avgLCL <- avgLCL
```

Update the BigDays.RData file. Create a back up version. 
```{r, eval = FALSE}
#save(All_Tornadoes, BigDayCentroids.df, BigDays.sfdfT, BigDayTornadoes, Groups.sfdfT, file = "BigDays.RData")
#save(All_Tornadoes, BigDayCentroids.df, BigDays.sfdfT, BigDayTornadoes, Groups.sfdfT, file = "BigDays_Backup.RData")
```
########################
## Population Density ##
########################

You need to establish a population density for each big day. Round the year to the nearest 10. Example: 1995 should round to 2000 census and 1993 should round to 1990 census. 

Create a function to round to the nearest 5 since we can get population for every 5 years starting 1995 - 2015. 
```{r}
mround <- function(x,base){ 
        base*round(x/base) 
} 
```

Add a new column `PopDensyear` to the dataset. This will round each year to the nearest 5th year. 
```{r, eval = FALSE}
BigDays.sfdfT <- BigDays.sfdfT %>%
  mutate(PopDensyear = mround(Year, 5))
```

Download the census data: 

Begin with The Gridded Population of the World Volume 3 data (2000). Begin with The Gridded Population of the World Volume 3 data. This data set has population density data available beginning in 1990 and ending in 2000. We are using the 1995 data. 

Population Data for 2000 through 2015 are obtained from the Gridded Population of the World, version four (GPW, v4) from the Socioeconomic Data and Applications Center at Columbia University, USA. The database contain decennial census density estimates for 2000, 2005, 2010, and 2015 represented as people per square kilometer. Densities are based on residential population. 

Load the population raster(s) and crop to defined extent. 
```{r, eval = FALSE}
#1995 Data: 
  rb <- raster("/Users/zoeschroder/Desktop/Projects/tornenvir_climate/PopDens/usadens/usads95g/w001001.adf")
  ext = raster::extent(c(-125, -67, 24, 50))
  PopDens = crop(rb, ext)
  PopDens1995 = projectRaster(PopDens, crs = "+proj=lcc +lat_1=50 +lat_2=50 +lat_0=50 +lon_0=-107 +x_0=0 +y_0=0 +a=6371200 +b=6371200 +units=m +no_defs")
#2000 Data: 
  rb <- raster(paste0("/Users/zoeschroder/Desktop/Projects/tornenvir_climate/PopDens", "/gpw-v4-population-density-rev11_2000_2pt5_min_tif/gpw_v4_population_density_rev11_2000_2pt5_min.tif"))
  ext = raster::extent(c(-125, -67, 24, 50))
  PopDens = crop(rb, ext)
  PopDens2000 = projectRaster(PopDens, crs = "+proj=lcc +lat_1=50 +lat_2=50 +lat_0=50 +lon_0=-107 +x_0=0 +y_0=0 +a=6371200 +b=6371200 +units=m +no_defs")
#2005 Data: 
    rb <- raster(paste0("/Users/zoeschroder/Desktop/Projects/tornenvir_climate/PopDens", "/gpw-v4-population-density-rev11_2005_2pt5_min_tif/gpw_v4_population_density_rev11_2005_2pt5_min.tif"))
  ext = raster::extent(c(-125, -67, 24, 50))
  PopDens = crop(rb, ext)
  PopDens2005 = projectRaster(PopDens, crs = "+proj=lcc +lat_1=50 +lat_2=50 +lat_0=50 +lon_0=-107 +x_0=0 +y_0=0 +a=6371200 +b=6371200 +units=m +no_defs")
#2010 Data: 
    rb <- raster(paste0("/Users/zoeschroder/Desktop/Projects/tornenvir_climate/PopDens", "/gpw-v4-population-density-rev11_2010_2pt5_min_tif/gpw_v4_population_density_rev11_2010_2pt5_min.tif"))
  ext = raster::extent(c(-125, -67, 24, 50))
  PopDens = crop(rb, ext)
  PopDens2010 = projectRaster(PopDens, crs = "+proj=lcc +lat_1=50 +lat_2=50 +lat_0=50 +lon_0=-107 +x_0=0 +y_0=0 +a=6371200 +b=6371200 +units=m +no_defs")
#2015 Data: 
    rb <- raster(paste0("/Users/zoeschroder/Desktop/Projects/tornenvir_climate/PopDens", "/gpw-v4-population-density-rev11_2015_2pt5_min_tif/gpw_v4_population_density_rev11_2015_2pt5_min.tif"))
  ext = raster::extent(c(-125, -67, 24, 50))
  PopDens = crop(rb, ext)
  PopDens2015 = projectRaster(PopDens, crs = "+proj=lcc +lat_1=50 +lat_2=50 +lat_0=50 +lon_0=-107 +x_0=0 +y_0=0 +a=6371200 +b=6371200 +units=m +no_defs")
#2020 Data: 
    rb <- raster(paste0("/Users/zoeschroder/Desktop/Projects/tornenvir_climate/PopDens", "/gpw-v4-population-density-rev11_2020_2pt5_min_tif/gpw_v4_population_density_rev11_2020_2pt5_min.tif"))
  ext = raster::extent(c(-125, -67, 24, 50))
  PopDens = crop(rb, ext)
  PopDens2015 = projectRaster(PopDens, crs = "+proj=lcc +lat_1=50 +lat_2=50 +lat_0=50 +lon_0=-107 +x_0=0 +y_0=0 +a=6371200 +b=6371200 +units=m +no_defs")
```

Use the `extract` function to obtain the population density.
```{r, eval = FALSE}
avgPopDens <- numeric()
maxPopDens <- numeric()
totalPopDens <- numeric()

library(raster)
for(i in 1:dim(BigDays.sfdfT)[1]){
  print(i)
  if (BigDays.sfdfT$PopDensyear[i] == 1995) {
  avgPopDens <- c(avgPopDens, as.numeric(raster::extract(PopDens1995, BigDays.sfdfT[i, ], fun = mean, na.rm = TRUE, weights = TRUE, normalizeWeights = FALSE)))
  maxPopDens <- c(maxPopDens, as.numeric(raster::extract(PopDens1995, BigDays.sfdfT[i, ], fun = max, na.rm = TRUE))) 
  totalPopDens <- c(totalPopDens, as.numeric(raster::extract(PopDens1995, BigDays.sfdfT[i, ], fun = sum, na.rm = TRUE))) 
  } else if (BigDays.sfdfT$PopDensyear[i] == 2000){
  avgPopDens <- c(avgPopDens, as.numeric(raster::extract(PopDens2000, BigDays.sfdfT[i,], fun = mean, na.rm = TRUE, weights = TRUE, normalizeWeights = FALSE)))
  maxPopDens <- c(maxPopDens, as.numeric(raster::extract(PopDens2000, BigDays.sfdfT[i,], fun = max, na.rm = TRUE)))
  totalPopDens <- c(totalPopDens, as.numeric(raster::extract(PopDens2000, BigDays.sfdfT[i, ], fun = sum, na.rm = TRUE))) 
  } else if (BigDays.sfdfT$PopDensyear[i] == 2005){
  avgPopDens <- c(avgPopDens, as.numeric(raster::extract(PopDens2005, BigDays.sfdfT[i,], fun = mean, na.rm = TRUE, weights = TRUE, normalizeWeights = FALSE)))
  maxPopDens <- c(maxPopDens, as.numeric(raster::extract(PopDens2005, BigDays.sfdfT[i,], fun = max, na.rm = TRUE)))
  totalPopDens <- c(totalPopDens, as.numeric(raster::extract(PopDens2005, BigDays.sfdfT[i, ], fun = sum, na.rm = TRUE))) 
  } else if (BigDays.sfdfT$PopDensyear[i] == 2010){
  avgPopDens <- c(avgPopDens, as.numeric(raster::extract(PopDens2010, BigDays.sfdfT[i,], fun = mean, na.rm = TRUE, weights = TRUE, normalizeWeights = FALSE)))
  maxPopDens <- c(maxPopDens, as.numeric(raster::extract(PopDens2010, BigDays.sfdfT[i,], fun = max, na.rm = TRUE)))
  totalPopDens <- c(totalPopDens, as.numeric(raster::extract(PopDens2010, BigDays.sfdfT[i, ], fun = sum, na.rm = TRUE))) 
  } else if (BigDays.sfdfT$PopDensyear[i] >= 2015){
  avgPopDens <- c(avgPopDens, as.numeric(raster::extract(PopDens2015, BigDays.sfdfT[i,], fun = mean, na.rm = TRUE, weights = TRUE, normalizeWeights = FALSE)))
  maxPopDens <- c(maxPopDens, as.numeric(raster::extract(PopDens2015, BigDays.sfdfT[i,], fun = max, na.rm = TRUE)))
  totalPopDens <- c(totalPopDens, as.numeric(raster::extract(PopDens2015, BigDays.sfdfT[i, ], fun = sum, na.rm = TRUE))) 
  } else if (BigDays.sfdfT$PopDensyear[i] == 2020){
  avgPopDens <- c(avgPopDens, as.numeric(raster::extract(PopDens2020, BigDays.sfdfT[i,], fun = mean, na.rm = TRUE, weights = TRUE, normalizeWeights = FALSE)))
  maxPopDens <- c(maxPopDens, as.numeric(raster::extract(PopDens2020, BigDays.sfdfT[i,], fun = max, na.rm = TRUE)))
  totalPopDens <- c(totalPopDens, as.numeric(raster::extract(PopDens2020, BigDays.sfdfT[i, ], fun = sum, na.rm = TRUE))) 
  }
}
```
 
Add the max and avg population density values to the Big Day data: 
```{r, eval = FALSE}
BigDays.sfdfT <- cbind(BigDays.sfdfT, avgPopDens) 
BigDays.sfdfT <- cbind(BigDays.sfdfT, maxPopDens) 
BigDays.sfdfT <- cbind(BigDays.sfdfT, totalPopDens) 
```

*Get the latitude and longitude for each big day centroid*
```{r, eval = FALSE}
BigDayCentroids.df <- st_centroid(BigDays.sfdfT)
test <- as_Spatial(BigDayCentroids.df)
spgeo <- spTransform(test, CRS("+proj=longlat +datum=WGS84"))
test <- st_as_sf(spgeo)
coords <- as.data.frame(st_coordinates(test))
colnames(coords)[colnames(coords)=="X"] <- "Lon"
colnames(coords)[colnames(coords)=="Y"] <- "Lat"

BigDays.sfdfT <- cbind(as.data.frame(BigDays.sfdfT), as.data.frame(coords))
```


```{r}
cities <- us_cities()
counties <- us_counties()
states <- us_states()
sts <- state.name[!state.name %in% c("Alaska", "Hawaii")]
stateBorders <- us_states(states = sts)
```

*Get the total population from the cities that fall within each big day*
```{r, eval = FALSE}
BigDays.sfdfT <- st_as_sf(BigDays.sfdfT)

BigDays.sfdfT <- st_convex_hull(BigDays.sfdfT)
BigDays.sfdfT <- st_transform(BigDays.sfdfT, 
  crs = "+proj=lcc +lat_1=50 +lat_2=50 +lat_0=50 +lon_0=-107 +x_0=0 +y_0=0 +a=6371200 +b=6371200 +units=m +no_defs")
cities <- st_transform(cities, 
  crs = "+proj=lcc +lat_1=50 +lat_2=50 +lat_0=50 +lon_0=-107 +x_0=0 +y_0=0 +a=6371200 +b=6371200 +units=m +no_defs")

totalPOP <- numeric()

for(i in 1:dim(BigDays.sfdfT)[1]){
    popinhull <- cities[BigDays.sfdfT[i,],]
    dayPOP <- sum(popinhull$population)
    totalPOP <- append(totalPOP, dayPOP)
}
BigDays.sfdfT <- cbind(BigDays.sfdfT, totalPOP)
```

Save the data.
```{r, eval = FALSE}
#save(All_Tornadoes, BigDayCentroids.df, BigDays.sfdfT, BigDayTornadoes, Groups.sfdfT, file = "BigDays.RData")
#save(All_Tornadoes, BigDayCentroids.df, BigDays.sfdfT, BigDayTornadoes, Groups.sfdfT, file = "BigDays_Backup.RData")
```

##################################
##Tornado Data has been updated.## 
##################################

Climate Data: 

```{r}
load("BigDays.RData")
```

Surface Temperatures from National Oceanic and Atmospheric Association's (NOAA) Global Surface Temperature (NOAAGlobalTemp) data provided by the NOAA/OAR/ESRL PSL (\url{https://psl.noaa.gov/data/gridded/data.noaaglobaltemp.html}). NOAAGlobalTemp is a combination of land and ocean temperatures combined into a singular dataset. The data includes monthly surface temperatures (both land and water) from 1880 to present. It has a spatial resolution of 5$^\circ$ by 5$^\circ$. 

In climate change studies, temperature anomalies are more important than absolute temperature. A temperature anomaly is the difference from an average, or baseline, temperature. The baseline temperature is typically computed by averaging 30 or more years of temperature data. A positive anomaly indicates the observed temperature was warmer than the baseline, while a negative anomaly indicates the observed temperature was cooler than the baseline. When calculating an average of absolute temperatures, things like station location or elevation will have an effect on the data (ex. higher elevations tend to be cooler than lower elevations and urban areas tend to be warmer than rural areas). However, when looking at anomalies, those factors are less critical. For example, a summer month over an area may be cooler than average, both at a mountain top and in a nearby valley, but the absolute temperatures will be quite different at the two locations.

Using anomalies also helps minimize problems when stations are added, removed, or missing from the monitoring network. The above diagram shows absolute temperatures (lines) for five neighboring stations, with the 2008 anomalies as symbols. Notice how all of the anomalies fit into a tiny range when compared to the absolute temperatures. Even if one station were removed from the record, the average anomaly would not change significantly, but the overall average temperature could change significantly depending on which station dropped out of the record. For example, if the coolest station (Mt. Mitchell) were removed from the record, the average absolute temperature would become significantly warmer. However, because its anomaly is similar to the neighboring stations, the average anomaly would change much less.

```{r}
library(raster)
ncpath <- "C:/Users/lordofmornin/Desktop/Projects/tornenvir_climate/"
ncname <- "air.mon.anom"  
ncfname <- paste(ncpath, ncname, ".nc", sep="")
library(ncdf4)
tmp_raster <- raster::brick(ncfname, varname="air")
class(tmp_raster)

January1994 <- subset(tmp_raster, 1369)
plot(January1994)

```

Transform the BigDays.sfdfT to the same projection as the raster data. Check the extent of BigDays.sfdfT to ensure the geometry is in degrees latitude and longitude. Rotate the `January1994` raster to get longitude in -180 to 180 degrees longitude (currently 0 to 360 degrees longitude).
```{r}
BigDays.sfdfT <- st_transform(BigDays.sfdfT, CRS("+proj=longlat +datum=WGS84"))

extent(BigDays.sfdfT)

January1994 <- rotate(January1994)
```

extent(BigDays.sfdfT)
class      : Extent 
xmin       : -116.93 
xmax       : -70.57 
ymin       : 24.9879 
ymax       : 49 

January1994
class      : RasterLayer 
dimensions : 36, 72, 2592  (nrow, ncol, ncell)
resolution : 5, 5  (x, y)
extent     : -180, 180, -90, 90  (xmin, xmax, ymin, ymax)
crs        : +proj=longlat +datum=WGS84 +no_defs 
source     : memory
names      : X1994.01.01 
values     : -8.546114, 5.79  (min, max)
z-value    : 1994-01-01 

#How do you set extent boundaries? Is it easier to manually set them? 
Create a spatial extent for the entire contiguous United States. You do this by creating a Spatial Polygons data frame then converting to an sf object. 

```{r}
USextent <- as(raster::extent(-130, -65, 25, 49), "SpatialPolygons") %>%
  st_as_sf()
```

Crop the January 1994 data to the extent of the United States. 
```{r}
Jan94 <- crop(January1994, extent(USextent))
Jan94_extent <- mask(Jan94, USextent)

plot(Jan94_extent)
```

Plot a map of the January 1994 air temperature data over the contiguous United States. 
```{r}
library(tmap)
tm_shape(Jan94_extent, 
         projection = merc, 
         is.master = TRUE, line.center.type = "midpoint") +
  tm_raster(title = "", midpoint = NA) +
  tm_format(title = expression(paste("January 1994 surface air temperature")),
                    "World", 
                    legend.position = c("right", "bottom"),
                    attr.position = c("right", "bottom"),
                    legend.frame = FALSE,
                    inner.margins = c(0, 0.01, 0, .01), #bottom, left, top
                    legend.text.size = 1, 
                    legend.width = -0.28, 
                    legend.title.size=1, 
                    legend.bg.color = "white", 
                    title.size = 1.4) +
#tm_shape(counties) +
#  tm_borders(col = "grey") +
  tm_scale_bar(width = 0.3, size = 1, position = c("right", "top"), color.dark = "gray70") +
tm_shape(stateBorders, midpoint = NA) +
  tm_borders() 
```






```{r}
test<- BigDays.sfdfT %>%
  filter(Year == 1994 & Mo %in% c(3,4))
```

```{r}
min_airtemp <- numeric()
avg_airtemp <- numeric()
max_airtemp <- numeric()

for (i in 1:dim(BigDays.sfdfT)[1]) {
  print(i)
  Year_add <- (BigDays.sfdfT$Year[i] - min(BigDays.sfdfT$Year)) * 12
  Month_add <- BigDays.sfdfT$Mo[i]
  total <- Year_add + Month_add
  val <- (1369 + total)
  ras <- subset(tmp_raster, val)
  ras <- rotate(ras)
  ras <- crop(ras, extent(USextent))
  ras <- mask(ras, USextent)
  min_airtemp <- c(min_airtemp, min(getValues(ras)))
  avg_airtemp <- c(avg_airtemp, mean(getValues(ras)))
  max_airtemp <- c(max_airtemp, max(getValues(ras)))
}
```

```{r}
BigDays.sfdfT$min_airtemp <- min_airtemp
BigDays.sfdfT$max_airtemp <- max_airtemp
BigDays.sfdfT$avg_airtemp <- avg_airtemp
```

Save the data.
```{r, eval = FALSE}
#save(All_Tornadoes, BigDayCentroids.df, BigDays.sfdfT, BigDayTornadoes, Groups.sfdfT, file = "BigDays.RData")
#save(All_Tornadoes, BigDayCentroids.df, BigDays.sfdfT, BigDayTornadoes, Groups.sfdfT, file = "BigDays_Backup.RData")
```


## Sea Surface Temperatures ##


PROSPECTUS: 15$^{\circ}$N to 25$^{\circ}$N and 70$^{\circ}$W to 90$^{\circ}$W

Sea surface temperatures for the western Caribbean. Latitude Range used: 20.0 to 4.8; Longitude Range used: 300.0 to 330.0
```{r}
#download.file(url = "https://psl.noaa.gov/cgi-bin/data/timeseries/timeseries.pl?ntype=1&var=SST&level=2000&lat1=15&lat2=25&lon1=270&lon2=290+&iseas=0&mon1=0&mon2=0&iarea=0&typeout=1&Submit=Create+Timeseries" ,
              #destfile = "Atlantic_SST.csv", mode = "wb")

Atl_SST <- read.csv(file = "Atlantic_SST.csv")

```

Get the sea surface temperature for each big day. 
```{r}
Atlantic_SST <- numeric()

for (i in 1:dim(BigDays.sfdfT)[1]) {
  print(i)
  month_column = as.numeric(BigDays.sfdfT$Month[i]) + 1
  year_row = BigDays.sfdfT$Year[i] - 1994
  Atlantic_SST <- c(Atlantic_SST, Atl_SST[47+ year_row, month_column]) #47 is year 1994 in the Atlantic_SST data frame
}  
```

Add sea surface temperatures to the BigDays.sfdfT file. 
```{r}
BigDays.sfdfT$Atlantic_SST <- Atlantic_SST
```

## Nino 3.4 ##
-99.99							
NINA34							
5N-5S 17	0W-120W						
HadISST							
https://	psl.noaa	.gov/g	cos_wgsp	/Timeser	ies/Nino	34/	

PROSPECTUS: 120$^{\circ}$W to 170$^{\circ}$W and 5$^{\circ}$N to 5$^{\circ}$S.

```{r}
download.file(url = "https://psl.noaa.gov/gcos_wgsp/Timeseries/Data/nino34.long.data",
              destfile = "Nino3_4.csv", mode = "wb")

NINO <- read.csv(file = "Nino3_4.csv")
```

```{r}
ElNino34 <- numeric()

for (i in 1:dim(BigDays.sfdfT)[1]) {
  print(i)
  month_column = as.numeric(BigDays.sfdfT$Month[i]) + 1
  year_row = BigDays.sfdfT$Year[i] - 1994
  ElNino34 <- c(ElNino34, NINO[125+ year_row, month_column]) #125 is year 1994 in the NINO data frame
}  
```

Add El Nino 3.4 data to the BigDays.sfdfT file. 
```{r}
BigDays.sfdfT$ElNino34 <- ElNino34
```


Arctic Sea Ice

Comes from the Sea Ice Index, Sea Ice Concentrations from Nimbus-7 SMMR SSMIS and DMSP SSM/I Passive Microwave Data is the input data that the Sea Ice Index is derived from. Extent and Area are in millions of square kilometers. N 90 to 30.98 E: 180 to -180 https://nsidc.org/data/g02135?qt-data_set_tabs=2#qt-data_set_tabs

```{r}
seaice <- read.csv(file = "SIE.csv")
```

```{r}
SIE <- numeric()

for (i in 1:dim(BigDays.sfdfT)[1]) {
  print(i)
  month_column = as.numeric(BigDays.sfdfT$Month[i]) * 2 + 2
  year_row = BigDays.sfdfT$Year[i] - 1994
  SIE <- c(SIE, seaice[16+ year_row, month_column]) #125 is year 1994 in the NINO data frame
}  
```

Add Sea Ice Extent to the BigDays.sfdfT file. 
```{r}
BigDays.sfdfT$SIE<- SIE
```

Save the data.
```{r, eval = FALSE}
#save(All_Tornadoes, BigDayCentroids.df, BigDays.sfdfT, BigDayTornadoes, Groups.sfdfT, file = "BigDays.RData")
#save(All_Tornadoes, BigDayCentroids.df, BigDays.sfdfT, BigDayTornadoes, Groups.sfdfT, file = "BigDays_Backup.RData")
```




